### 2023-10-11 docs/manuscript/main.tex
added template and draft for ilp

\documentclass{article}[7pt]
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\include{ilputil}
\title{spp-dcj-cf}
\author{lbohnenk }
\date{June 2023}
\usepackage{algorithm}
\usepackage{placeins}
\usepackage{bbm}
\begin{document}
\newcommand{\extremityedges}{E_{ext}}
\newcommand{\adjacencyedges}{E_{adj}}
\newcommand{\indeledges}{E_{id}}
\newcommand{\idx}[1]{\texttt{ix}(#1)}
\newcommand{\parityvar}{b}
\newcommand{\caps}{V_{\varnothing}}
\newcommand{\trans}{|}
\newcommand{\cis}{\circ}
\newcommand{\coolone}{\mathbbm{1}}

\begin{algorithm}[htb]
\caption{new ILP}
\textbf{Minimize:}
\begin{equation*}
    n - c -q
\end{equation*}

\begin{constraints}

\begin{tabular}{lcl}
    \cns & $x_{e} = 1 $ & $\forall e\in \adjacencyedges$\\
     \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= 1$ & $\forall u\in V\setminus V_{\varnothing}$  \\
     \cns & $x_e=x_d$ & for all sibling edges $e,d$\\
     \cns & $y_v + \idx{u}(1-x_{uv}) \geq y_u$ &$\forall uv \in \adjacencyedges\cup \extremityedges$\\
          & $\idx{u}(1-x_{uv})\geq y_u$& $\forall uv \in \indeledges$\\
    \cns & $\idx{v}z_v \leq y_v$ & $\forall v\in V $\\
    \cns & matching condition\\
    \cns & $\parityvar_v = \coolone_{v \text{ of genome } B}$ & $\forall v \in \caps$\\
    \cns & $\parityvar_v \leq 1 - x_{uv}$ & $\forall uv \in \indeledges \text{ of genome } A$\\
     & $\parityvar_v \geq  x_{uv}$ & $\forall uv \in \indeledges \text{ of genome } B$\\
    \cns & $\parityvar_v \leq \parityvar_u +  x_{uv}$& $\forall uv \in \extremityedges$\\
    & $\parityvar_u \leq \parityvar_v + r_{uv}^{a\trans b} + (1-x_{uv})$& $\forall uv\in E_{adj},u,v\notin \caps$\\
    & $\parityvar_u \leq \parityvar_v + r_{uv}^{A\trans B} + r_{uv}^{A\trans b} + (1-x_{uv})$& $\forall uv\in E_{adj}, u \text{ or } v\in \caps^A$\\
    & $\parityvar_u \leq \parityvar_v + r_{uv}^{B\trans a} + (1-x_{uv})$& $\forall uv\in E_{adj}, u \text{ or } v\in \caps^B$\\
    \cns & $a_v = 0$ & $\forall v\in V_{\varnothing}$\\
    \cns & $x_{uv} \leq a_v$&$\forall uv\in E_{id}$\\
    \cns & $a_v \leq a_u + (1-x_{uv})$&$\forall uv\in E_{ext}$\\
    \cns & $a_v \leq a_u + r_{uv}^{A\trans b} + r_{uv}^{A\cis a}$&$\forall uv\in E_{adj}, u \text{ or } v\in \caps^A$\\
    & $a_v \leq a_u + r_{uv}^{B\trans a} + r_{uv}^{B\cis b}$&$\forall uv\in E_{adj}, u \text{ or } v\in \caps^B$\\
    \cns & $r_{uv}^c + r_{uv}^{A\trans B} \leq z_v + z_u$&$\forall uv \in E_{adj}$\\
    \cns & $r_{uv}^{R} \leq b_u + b_v$\\
    \cns & $r_{uv}^{R} \leq 2-b_u - b_v$&$\forall R \in \{A\trans B, A\trans b, B\trans a\}$\\
    \cns & $r_{uv}^{a\trans b} \leq x_{uw} + x_{vz}$&$\forall uv\in E_{adj}, uw,vz \in E_{ide}$\\
    \cns & $(u+v)(1-r_{uv}^{P}) \geq y_u +y_v$&$R\in  \{A\trans b , B\trans a , a\trans b, A\cis a , B\cis b\}$\\
    \cns & $ r_{uv}^{c\cis}\leq a_v + a_u $\\

    \cns & $n = 0.5 \sum_{e \in E_{ext}} x_{e}$\\
    \cns & $c = \sum_{uv\in E_{adj}} r^c_{uv}$\\
    \cns & $2q \geq p^{a\trans b}+ p^{ABa} + p^{ABb} - p^{A\trans B} $\\
    \cns & $p^{a\trans b} = \sum_{uv \in E_{adj}}r^{a\trans b}_{uv}$\\
    \cns & $p^{A\trans b} = \sum_{uv \in E_{adj}} r^{A\trans b}$\\
    \cns & $p^{B\trans a} = \sum_{uv \in E_{adj}} r^{B\trans a}$\\
    \cns & $p^{A\cis a} = \sum_{uv \in E_{adj}} r^{A\cis a}$\\
    \cns & $p^{B\cis b} = \sum_{uv \in E_{adj}} r^{B\cis b}$\\
    \cns & $p^{ABa} \geq p^{B\trans a}$\\
    \cns & $p^{ABa} \geq p^{A\cis a}$\\
    \cns & $p^{ABb} \geq p^{B\cis b}$\\
    \cns & $p^{ABb} \geq p^{A\trans b}$\\
    \cns & $p^{A\trans B} = \sum_{uv\in E_{adj}} r^{A\trans B}$\\
    \cns\cns & Circ sing\\
    \cns & $r^{c}_{uv} \leq x_{uv}$& only for spp-DCJ\\
\end{tabular}
\end{constraints}

\end{algorithm}
\end{document}

### 2023-10-19 docs/manuscript/main.tex
add comment w.r.t. matchings
Add some initial commands

@@
-\documentclass{article}[7pt]
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\include{ilputil}
-\title{spp-dcj-cf}
-\author{lbohnenk }
-\date{June 2023}
-\usepackage{algorithm}
-\usepackage{placeins}
-\usepackage{bbm}
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.21 of 2022/01/12
+%
+\documentclass[runningheads]{llncs}
+%
+\input{preamble}
+\usepackage[T1]{fontenc}
+% T1 fonts will be used to generate the final print and online PDFs,
+% so please use T1 fonts in your manuscript whenever possible.
+% Other font encondings may result in incorrect characters.
+%
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following two lines
+% to display URLs in blue roman font according to Springer's eBook style:
+
+%\renewcommand\UrlFont{\color{blue}\rmfamily}
+%
+
 \begin{document}
-\newcommand{\extremityedges}{E_{ext}}
-\newcommand{\adjacencyedges}{E_{adj}}
-\newcommand{\indeledges}{E_{id}}
-\newcommand{\idx}[1]{\texttt{ix}(#1)}
-\newcommand{\parityvar}{b}
-\newcommand{\caps}{V_{\varnothing}}
-\newcommand{\trans}{|}
-\newcommand{\cis}{\circ}
-\newcommand{\coolone}{\mathbbm{1}}
+%
+\title{Reconstructing Rearrangement Phylogenies}
+%
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
+Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
+Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
+%
+\authorrunning{F. Author et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Princeton University, Princeton NJ 08544, USA \and
+Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
+\email{lncs@springer.com}\\
+\url{http://www.springer.com/gp/computer-science/lncs} \and
+ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
+\email{\{abc,lncs\}@uni-heidelberg.de}}
+%
+\maketitle              % typeset the header of the contribution
+%
+\begin{abstract}
+The abstract should briefly summarize the contents of the paper in
+150--250 words.
 
-\begin{algorithm}[htb]
-\caption{new ILP}
-\textbf{Minimize:}
-\begin{equation*}
-    n - c -q
-\end{equation*}
+\keywords{First keyword  \and Second keyword \and Another keyword.}
+\end{abstract}
+%
+%
+%
 
-\begin{constraints}
+\section{Introduction}
+\section{Background}
+\subsection{Problem Formulation}
+\section{A new Method}
+\subsection{Pre-selecting Adjacencies}
+\todoDany{Describe Process}
+\subsection{A New ILP Formulation}
+\todoLeo{ILP goes here}
 
-\begin{tabular}{lcl}
-    \cns & $x_{e} = 1 $ & $\forall e\in \adjacencyedges$\\
-     \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= 1$ & $\forall u\in V\setminus V_{\varnothing}$  \\
-     \cns & $x_e=x_d$ & for all sibling edges $e,d$\\
-     \cns & $y_v + \idx{u}(1-x_{uv}) \geq y_u$ &$\forall uv \in \adjacencyedges\cup \extremityedges$\\
-          & $\idx{u}(1-x_{uv})\geq y_u$& $\forall uv \in \indeledges$\\
-    \cns & $\idx{v}z_v \leq y_v$ & $\forall v\in V $\\
-    \cns & matching condition\\
-    \cns & $\parityvar_v = \coolone_{v \text{ of genome } B}$ & $\forall v \in \caps$\\
-    \cns & $\parityvar_v \leq 1 - x_{uv}$ & $\forall uv \in \indeledges \text{ of genome } A$\\
-     & $\parityvar_v \geq  x_{uv}$ & $\forall uv \in \indeledges \text{ of genome } B$\\
-    \cns & $\parityvar_v \leq \parityvar_u +  x_{uv}$& $\forall uv \in \extremityedges$\\
-    & $\parityvar_u \leq \parityvar_v + r_{uv}^{a\trans b} + (1-x_{uv})$& $\forall uv\in E_{adj},u,v\notin \caps$\\
-    & $\parityvar_u \leq \parityvar_v + r_{uv}^{A\trans B} + r_{uv}^{A\trans b} + (1-x_{uv})$& $\forall uv\in E_{adj}, u \text{ or } v\in \caps^A$\\
-    & $\parityvar_u \leq \parityvar_v + r_{uv}^{B\trans a} + (1-x_{uv})$& $\forall uv\in E_{adj}, u \text{ or } v\in \caps^B$\\
-    \cns & $a_v = 0$ & $\forall v\in V_{\varnothing}$\\
-    \cns & $x_{uv} \leq a_v$&$\forall uv\in E_{id}$\\
-    \cns & $a_v \leq a_u + (1-x_{uv})$&$\forall uv\in E_{ext}$\\
-    \cns & $a_v \leq a_u + r_{uv}^{A\trans b} + r_{uv}^{A\cis a}$&$\forall uv\in E_{adj}, u \text{ or } v\in \caps^A$\\
-    & $a_v \leq a_u + r_{uv}^{B\trans a} + r_{uv}^{B\cis b}$&$\forall uv\in E_{adj}, u \text{ or } v\in \caps^B$\\
-    \cns & $r_{uv}^c + r_{uv}^{A\trans B} \leq z_v + z_u$&$\forall uv \in E_{adj}$\\
-    \cns & $r_{uv}^{R} \leq b_u + b_v$\\
-    \cns & $r_{uv}^{R} \leq 2-b_u - b_v$&$\forall R \in \{A\trans B, A\trans b, B\trans a\}$\\
-    \cns & $r_{uv}^{a\trans b} \leq x_{uw} + x_{vz}$&$\forall uv\in E_{adj}, uw,vz \in E_{ide}$\\
-    \cns & $(u+v)(1-r_{uv}^{P}) \geq y_u +y_v$&$R\in  \{A\trans b , B\trans a , a\trans b, A\cis a , B\cis b\}$\\
-    \cns & $ r_{uv}^{c\cis}\leq a_v + a_u $\\
+\section{Evaluation}
+\section{Discussion}
 
-    \cns & $n = 0.5 \sum_{e \in E_{ext}} x_{e}$\\
-    \cns & $c = \sum_{uv\in E_{adj}} r^c_{uv}$\\
-    \cns & $2q \geq p^{a\trans b}+ p^{ABa} + p^{ABb} - p^{A\trans B} $\\
-    \cns & $p^{a\trans b} = \sum_{uv \in E_{adj}}r^{a\trans b}_{uv}$\\
-    \cns & $p^{A\trans b} = \sum_{uv \in E_{adj}} r^{A\trans b}$\\
-    \cns & $p^{B\trans a} = \sum_{uv \in E_{adj}} r^{B\trans a}$\\
-    \cns & $p^{A\cis a} = \sum_{uv \in E_{adj}} r^{A\cis a}$\\
-    \cns & $p^{B\cis b} = \sum_{uv \in E_{adj}} r^{B\cis b}$\\
-    \cns & $p^{ABa} \geq p^{B\trans a}$\\
-    \cns & $p^{ABa} \geq p^{A\cis a}$\\
-    \cns & $p^{ABb} \geq p^{B\cis b}$\\
-    \cns & $p^{ABb} \geq p^{A\trans b}$\\
-    \cns & $p^{A\trans B} = \sum_{uv\in E_{adj}} r^{A\trans B}$\\
-    \cns\cns & Circ sing\\
-    \cns & $r^{c}_{uv} \leq x_{uv}$& only for spp-DCJ\\
-\end{tabular}
-\end{constraints}
+\nocite{DOE-CHA-2021}
+\nocite{BOH-BRA-DOE-STO-2021}
+\subsubsection{Acknowledgements} Please place your acknowledgments at
+the end of the paper, preceded by an unnumbered run-in heading (i.e.
+3rd-level heading).
 
-\end{algorithm}
+%
+% ---- Bibliography ----
+%
+% BibTeX users should specify bibliography style 'splncs04'.
+% References will then be sorted and formatted in the correct style.
+%
+\bibliographystyle{splncs04}
+\bibliography{refs}
+
 \end{document}

### 2023-10-19 docs/manuscript/main.tex
Include ILP into text. TODO: Descriptions of the ILP, fix inconsistencies of variables.

@@
 %
-\title{Reconstructing Rearrangement Phylogenies}
+\title{\todoDany{Approve or change: } Reconstructing Rearrangement Phylogenies}
 %
@@
 %
+\todo{Authors and order? $\rightarrow$ until Oct 23rd.}
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+\todo{We have to submit this until Oct 23rd.}
 
@@
 \todoLeo{ILP goes here}
+\todoLeo{Remove idx stuff and bring back $v_i$ notation}
+
+\begin{algorithm}
+\caption{Capping-free Small Parsimony}
+\textbf{Objective}
+
+\newcommand{\idx}{\texttt{ix}}
+\hspace{0.5cm}\texttt{Maximize}
+\begin{equation*}
+    \alpha \sum_{\edt\in\tree} f_\edt + .... \text{\todoDany{Optimization of other factors goes here.}}
+\end{equation*}
+
+\textbf{For each tree edge $(\A,\B):=\edt\in \tree$ with $\pcmrd(\A,\B) = (\vertices\cup\pseudocaps, \adjacencyedges \cup \extremityedges \cup \selfedges)$:}
+
+\begin{constraints}
+\begin{tabular}{lcl}
+	\cns & $f_\edt = n_\edt - c_\edt + q_\edt + s_\edt$\\
+	\cns & $n_\edt = 0.5 \sum_{e \in \extremityedges} x_{e}$\\
+    \cns & $c_\edt = \sum_{v\in \vertices} r^c_{v}$\\
+    \cns & $2q_\edt \geq p^{a\trans b}+ p^{ABa} + p^{ABb} - p^{A\trans B} $\\
+    \cns & $p^{a\trans b} = \sum_{v \in \vertices}r^{a\trans b}_{v}$\\
+    \cns & $p^{A\trans b} = \sum_{v \in \pseudocaps^A} r^{A\trans b}_e$\\
+    \cns & $p^{B\trans a} = \sum_{v\in \pseudocaps^B} r^{B\trans a}_e$\\
+    \cns & $p^{A\cis a} = \sum_{v \in \pseudocaps^A} r^{A\cis a}_e$\\
+    \cns & $p^{B\cis b} = \sum_{e \in \pseudocaps^B} r^{B\cis b}_e$\\
+    \cns & $p^{ABa} \geq p^{B\trans a}$\\
+    \cns & $p^{ABa} \geq p^{A\cis a}$\\
+    \cns & $p^{ABb} \geq p^{B\cis b}$\\
+    \cns & $p^{ABb} \geq p^{A\trans b}$\\
+    \cns & $p^{A\trans B} = \sum_{v\in \pseudocaps^A} r^{A\trans B}_v$\\
+    \cns & $s_\edt = \sum_{v\in\vertices} r^{s}_{v}$ \\
+    (\ref{ilp:slmstart}) to (\ref{ilp:slmend})& Shao-Lin-Moret~\cite{SHA-LIN-MOR-2015} constraints& -- see Table~\ref{tab:slmcons}\\
+    (\ref{ilp:regvstart}) to (\ref{ilp:regvend})& Reporting for regular vertices& -- see Table~\ref{tab:regv}\\
+    (\ref{ilp:pcstart}) to (\ref{ilp:pcend})& Reporting for \gls{pcaps}& -- see Table~\ref{tab:pcaps}\\
+    (\ref{ilp:csstart}) to (\ref{ilp:csend})& Reporting circular singletons& --  see Table~\ref{tab:csreport}\\    
+\end{tabular}
+
+\end{constraints}
+\todoLeo{Put domains here?}
+\todoLeo{Fix formatting}
+\end{algorithm}
+
+\newcommand{\idx}{}
+\begin{table}
+
+\begin{constraints}
+\caption{Shao-Lin-Moret constraints.} \label{tab:slmcons}
+\begin{tabular}{lcl}
+ \cns\label{ilp:slmstart} & $\sum_{uv\in \ades} x_{e} = 1 $ & $\forall u\in \vertices$\\
+     \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= 1$ & $\forall u\in \vertices$  \\
+     \cns & $x_e=x_d$ & for all sibling edges $e,d$\\
+     \cns & $y_v + \idx{u}(1-x_{uv}) \geq y_u$ &$\forall uv \in \adjacencyedges\cup \extremityedges$\\
+          & $\idx{u}(1-x_{uv})\geq y_u$& $\forall uv \in \selfedges$\\
+    \cns\label{ilp:slmend} & $\idx{v}z_v \leq y_v$ & $\forall v\in \vertices\cup\pseudocaps $\\
+\end{tabular}
+\end{constraints}
+
+\end{table}
+
+\begin{table}
+
+\begin{constraints}
+\caption{Reporting for regular vertices.\comment{This is basically exactly the same as in ding, just swap out $r^{ab}$ for $t$ and $\parityvar$ for $r$ and root reports on vertices instead of edges.}} \label{tab:regv}
+\begin{tabular}{lcl}
+\cns\label{ilp:regvstart} & $\parityvar^\edt_v \leq 1 - x_{uv}$ & $\forall uv \in \selfedges^\A$\\
+     & $\parityvar^\edt_v \geq  x_{uv}$ & $\forall uv \in \selfedges^\B$\\
+\cns & $\parityvar_v \leq \parityvar_u +  (1-x_{uv})$& $\forall uv \in \extremityedges$\\
+& $\parityvar_u \leq \parityvar_v + r_{uv}^{a\trans b} + (1-x_{uv})$& $\forall uv\in E_{adj}^\A,u,v\notin \pseudocaps$\\
+& $\parityvar_u \leq \parityvar_v + (1-x_{uv})$& $\forall uv\in E_{adj}^\B,u,v\notin \pseudocaps$\\
+
+\cns & $r_{v}^c \leq z_v$&$\forall v \in \vertices^\A$\\
+\cns\label{ilp:regvend} & $r_{u}^{a\trans b} \leq x_{uv}$&$\forall uv\in\selfedges^\A$\\
+\end{tabular}
+\end{constraints}
+
+\end{table}
+
+\begin{table}
+
+\begin{constraints}
+\caption{Reporting for \gls{pcaps}.} \label{tab:pcaps}
+\begin{tabular}{lcl}
+\cns\label{ilp:pcstart} & $\parityvar_v = 0$ & $\forall v \in \pseudocaps^\A$\\
+ & $\parityvar_v = 1$ & $\forall v \in \pseudocaps^\B$\\
+\cns & $\parityvar_u \leq \parityvar_v + r_{v}^{A\trans B} + r_{v}^{A\trans b} + (1-x_{uv})$& $\forall uv\in \ades, v \in \pseudocaps^\A$\\
+    & $\parityvar_u \leq \parityvar_v + r_{u}^{B\trans a} + (1-x_{uv})$& $\forall uv\in \ades, u\in \pseudocaps^\B$\\
+    \cns & $r_{v}^{A\trans B} \leq z_v$&$\forall v \in \pseudocaps^\A$\\
+    \cns & $1 -y_v \leq r^{A\trans b}_{v} + r^{A\cis a}_v$ & $v \in \pseudocaps^\A$\\
+        & $1 -y_v \leq r^{B\trans a}_{v} + r^{B\cis b}_v$ & $v \in \pseudocaps^\B$\\
+    \cns & $y_{v_i} \leq i(1-r_{v}^{R})$&$v\in \pseudocaps^\A,R\in  \{A\trans b, A\cis a\}$\\
+     & $y_{v_i} \leq i(1-r_{v}^{R})$&$v\in \pseudocaps^\B,R\in  \{B\trans a, B\cis b\}$\\
+    \cns\label{ilp:pcend} & $r_{v}^{A\trans B} \leq b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^A$ \\
+    & $r_{v}^{A\trans b} \leq b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^A$ \\
+    & $r_{v}^{B\trans a} \leq 1-b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^\B$\\
+\end{tabular}
+\end{constraints}
+
+\end{table}
+
+\begin{table} \caption{Reporting circular singletons.\comment{(This is what we developed last week.)}}\label{tab:csreport}
+\begin{tabular}{lcl}
+     \cns\label{ilp:csstart} & $d_u+d_v + x_{uv}\leq 2$ & $\forall uv \in \adjacencyedges\cup \selfedges$  \\
+     & $d_u + d_v - x_{uv} \geq 0$ & $\forall uv \in \adjacencyedges\cup \selfedges$  \\
+    \cns & $w_u = w_v$ & $\forall uv \in \selfedges$\\
+    \cns\label{ilp:csend} &$K (1- x_{uv} + r^{s}_{u} + r^{s}_{v}) + w_v \geq w_u + d_v - d_u $ & $\forall uv\in \adjacencyedges$\\
+     %& $K (1- x_{uv} + r^{s}_{v} + r^{s}_{u}) + w_u \geq w_v + d_u - d_v $ & $\forall uv\in \adjacencyedges$\\
+\end{tabular}
+\end{table}
+
+\todoLeo{Fix inconsistencies with comparison in $\parityvar$ and $z$.}
 
@@
 
-\nocite{DOE-CHA-2021}
-\nocite{BOH-BRA-DOE-STO-2021}
 \subsubsection{Acknowledgements} Please place your acknowledgments at

### 2023-10-20 docs/manuscript/main.tex
committing intermediate state

@@
 %
-\title{\todoDany{Approve or change: } Reconstructing Rearrangement Phylogenies}
+\title{Reconstructing Rearrangement Phylogenies of Natural Genomes}
 %
@@
 %
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
+\author{Leonard Bohnen\"amper\inst{1}\orcidID{0000-0003-4508-0078} \and
+    Jens Stoye \inst{1}\orcidID{0000-0002-4656-7155} \and
+    Daniel D\"orr \inst{2}\orcidID{0000-0002-3720-6227}
+}
 %
-\authorrunning{F. Author et al.}
+\authorrunning{L. Bohnenk\"amper et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
+\institute{Faculty of Technology and Center for Biotechnology (CeBiTec), Bielefeld University, Germany\\ 
+    \email{lbohnenkaemper@techfak.uni-bielefeld.de}, \email{stoye@techfak.uni-bielefeld.de}\\
+\and
+Institute for Medical Biometry and Bioinformatics, Medical Faculty, and Center for Digital Medicine, Heinrich Heine University, Germany\\
+\email{daniel.doerr@hhu.de}}
+
 %
@@
 %
-\todo{Authors and order? $\rightarrow$ until Oct 23rd.}
 \begin{abstract}
-\todo{We have to submit this until Oct 23rd.}
+    We study the classic problem of inferring ancestral genomes on a given phylogeny and a given set of extant genomes, also known as \emph{small parsimony problem} (SPP). 
+    More specifically, we address a scenario where genomes represent sets of linear or circular successions of oriented genomic markers orders, also known as \emph{gene orders} that evolve along the branches of the phylogeny and are altered by a general evolutionary model that supports a genome rearrangements emulated through the double-cut-and-join (DCJ) operation and by segmental insertions and deletions.
+
+    Even under simple evolutionary models, this problem is computationally intractable,
 

### 2023-10-20 docs/manuscript/main.tex
added global constraints

@@
 
+\textbf{For each ancestral genome $\X \in \tree$ with $\ggraph(\X) = (\vertices, \adjacencyedges)$}
+
+\begin{constraints}
+\begin{tabular}{lcl}
+    \cns & $\minfam{F} \leq \sum_{v \in F} g_v \leq \maxfam{F}$ & $\forall F \in \families$\\
+    \cns & $g_v = g_u$ & for each pair of nodes $(u, v)$ corresponding to head/tail extremity of the same gene\todoDany{improve}\\
+    \cns & $g_{v} = x^{\X\A}_{v} = x^{\X\B}_{v} = x^{\X\C}_{v}$& $\forall v \in V$ and genomes $\A$, $\B$, $\C$ adjacent to $\X$ in $\Gamma$\\
+    \cns & $\sum_{uv \in \adjacencyedges} a_{uv} = g_u$ & $\forall u \in \vertices$\\
+    \cns & $a_{uv} = x^{\X\A}_{uv} = x^{\X\B}_{uv} = x^{\X\C}_{uv}$& $\forall uv \in \adjacencyedges$ and genomes $\A$, $\B$, $\C$ adjacent to $\X$ in $\Gamma$\\
+\end{tabular}
+\end{constraints}
+
+\medskip
 \textbf{For each tree edge $(\A,\B):=\edt\in \tree$ with $\pcmrd(\A,\B) = (\vertices\cup\pseudocaps, \adjacencyedges \cup \extremityedges \cup \selfedges)$:}
@@
     \cns & $s_\edt = \sum_{v\in\vertices} r^{s}_{v}$ \\
+     \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= x_u$ & $\forall u\in \vertices$  \\
     (\ref{ilp:slmstart}) to (\ref{ilp:slmend})& Shao-Lin-Moret~\cite{SHA-LIN-MOR-2015} constraints& -- see Table~\ref{tab:slmcons}\\
@@
 \begin{tabular}{lcl}
- \cns\label{ilp:slmstart} & $\sum_{uv\in \ades} x_{e} = 1 $ & $\forall u\in \vertices$\\
-     \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= 1$ & $\forall u\in \vertices$  \\
-     \cns & $x_e=x_d$ & for all sibling edges $e,d$\\
+% \cns\label{ilp:slmstart} & $\sum_{uv\in \ades} x_{e} = 1 $ & $\forall u\in \vertices$\\
+    \cns\label{ilp:slmstart} & $x_e=x_d$ & for all sibling edges $e,d$\\
     \cns & $y_v + \idx{u}(1-x_{uv}) \geq y_u$ &$\forall uv \in \adjacencyedges\cup \extremityedges$\\

### 2023-10-20 docs/manuscript/main.tex
implemented Leonards mapping idea

@@
     \cns & $g_v = g_u$ & for each pair of nodes $(u, v)$ corresponding to head/tail extremity of the same gene\todoDany{improve}\\
-    \cns & $g_{v} = x^{\X\A}_{v} = x^{\X\B}_{v} = x^{\X\C}_{v}$& $\forall v \in V$ and genomes $\A$, $\B$, $\C$ adjacent to $\X$ in $\Gamma$\\
+%    \cns & $g_{v} = x^{\X\A}_{v} = x^{\X\B}_{v} = x^{\X\C}_{v}$& $\forall v \in V$ and genomes $\A$, $\B$, $\C$ adjacent to $\X$ in $\Gamma$\\
     \cns & $\sum_{uv \in \adjacencyedges} a_{uv} = g_u$ & $\forall u \in \vertices$\\
-    \cns & $a_{uv} = x^{\X\A}_{uv} = x^{\X\B}_{uv} = x^{\X\C}_{uv}$& $\forall uv \in \adjacencyedges$ and genomes $\A$, $\B$, $\C$ adjacent to $\X$ in $\Gamma$\\
 \end{tabular}
@@
 \begin{tabular}{lcl}
+
 	\cns & $f_\edt = n_\edt - c_\edt + q_\edt + s_\edt$\\
@@
     \cns & $s_\edt = \sum_{v\in\vertices} r^{s}_{v}$ \\
-     \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= x_u$ & $\forall u\in \vertices$  \\
+    \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= g_{\map(u)}$ & $\forall u\in \vertices$  \\
+    \cns & $a_{\map(u)\map(v)} = x_{uv}$ & $\forall uv \in \adjacencyedges$ \\
+
     (\ref{ilp:slmstart}) to (\ref{ilp:slmend})& Shao-Lin-Moret~\cite{SHA-LIN-MOR-2015} constraints& -- see Table~\ref{tab:slmcons}\\

### 2023-10-21 scripts/spp_dcj.py
Revert "Various fixes. ILP still untested, but at least solvable."
This reverts commit 90a15f7b268e940512d3d995e8a9e67528dc6b1b.

@@ -31,7 +31,7 @@ LOG.setLevel(logging.DEBUG)
 # ILP OBJECTIVE
 #
 
-def objective(graphs, circ_singletons, alpha, beta, out,fullform=False):
+def objective(graphs, circ_singletons, alpha, beta, out):
     out.write('maximize ')
 
     # sum of adjacency weights over all genomes
@@ -64,8 +64,6 @@ def objective(graphs, circ_singletons, alpha, beta, out,fullform=False):
             x[2]['type'] == du.ETYPE_ADJ, G.edges(data = True)))) for i, (_, G) in \
             enumerate(sorted(graphs.items()))))):
         out.write(f' - {weight * beta} x{adj}')
-    if fullform:
-        write_missing_formula(graphs,alpha,out)
 
     out.write('\n\n')
 
@@ -339,15 +337,9 @@ def cfc03(siblings,out):
     return c03(siblings,out)
 
 
-def cfc04(G,i,out):
-    for u, v, data in G.edges(data = True):
-        if data['type']==du.ETYPE_ID:
-            #indel edges only exist 'virtually'
-            continue
-        out.write('y{0}_{3} - y{1}_{3} + {0} x{2} <= {0}\n'.format(u, v,
-            data['id'], i))
-        out.write('y{1}_{3} - y{0}_{3} + {1} x{2} <= {1}\n'.format(u, v,
-            data['id'], i))
+def cfc04(G,compnum,out):
+    #again, no difference between cf and capped version
+    return c04(G,compnum,out)
 
 def cfc05(G,compnum,out):
     #again, no difference between cf and capped version
@@ -372,7 +364,7 @@ def cfc06(G,compnum,out,genomes):
             sptype = TELOMERE_A
         else:
             sptype = TELOMERE_B
-        line='t{} - {} <= 0\n'.format(v,nvar(mrd=compnum,tp=sptype,v=v))
+        line='t_{} - {} <= 0\n'.format(v,mvar(mrd=compnum,tp=sptype,v=v))
         out.write(line)
 
 def cfc08(G,compnum,out):
@@ -468,7 +460,7 @@ def cfc17(circ_singletons, i, out):
 
 
 
-def cf_objective(graphs,circ_singletons,alpha,beta,out,fullform=False):
+def cf_objective(graphs,circ_singletons,alpha,beta,out):
     #TODO: What about telomeric, ie. artificial adjacencies?
     #How to get their weight?
[…]

### 2023-10-21 scripts/spp_dcj.py
Revert "Fix setting of z variable. Major bugs remain."
This reverts commit 669b0abfbfa82db004faf8e5d405eb47073699ca.

@@ -343,7 +343,7 @@ def cfc04(G,compnum,out):
 
 def cfc05(G,compnum,out):
     #again, no difference between cf and capped version
-    return c06(G,compnum,out)
+    return c05(G,compnum,out)
 
 def cfc07(G,compnum,out,genomes):
     for v in G.nodes():
@@ -414,7 +414,7 @@ def cfc12(G,compnum,out):
         for ij in PATHTYPES:
             if ij==PTYPE_CYCLE:
                 line = ' + '.join([mvar(mrd=compnum,tp=tp,v=v) for tp in SUBPATHTYPES])
-                line+= ' + ' +( ' + '.join([nvar(mrd=compnum,tp=tp,v=v) for tp in SUBPATHTYPES]))
+                line+= ' + '.join([nvar(mrd=compnum,tp=tp,v=v) for tp in SUBPATHTYPES])
                 line+= ' + 8 {} <= 8\n'.format(reportvar(tp=PTYPE_CYCLE,mrd=compnum,v=v))
                 out.write(line)
             else:

### 2023-10-21 scripts/spp_dcj.py
Revert "Various minor fixes."
This reverts commit c6e7115dae7dbbf6491bc9e04ab62840c01962eb.

@@ -446,11 +446,11 @@ def cfc16(G,compnum,out):
     if numerator:
         numerator+=' + '
     numerator+=summationvar(qr='r'+INDEL_A,mrd=compnum)
-    numerator+=' + '+summationvar(qr='r'+INDEL_B,mrd=compnum)
+    numerator+=summationvar(qr='r'+INDEL_B,mrd=compnum)
     negative = ' - '.join([reportvar(mrd=compnum,v=v,tp=''.join(sorted([TELOMERE_A,TELOMERE_B]))) for v in G.nodes()])
     if negative:
         numerator+=(" - "+negative)
-    line = numerator+" - 2 {} <= 0\n".format(summationvar(qr='q',mrd=compnum))
+    line = numerator+" - {}\n".format(summationvar(qr='q',mrd=compnum))
     out.write(line)
 
 def cfc17(circ_singletons, i, out):
@@ -463,7 +463,6 @@ def cfc17(circ_singletons, i, out):
 def cf_objective(graphs,circ_singletons,alpha,beta,out):
     #TODO: What about telomeric, ie. artificial adjacencies?
     #How to get their weight?
-    out.write('maximize ')
     written=False
     adjs = set(reduce(lambda x, y: x + y, (tuple(map(lambda z: (z[2]['id'], \
             z[2]['weight']), filter(lambda x: x[2]['type'] == du.ETYPE_ADJ, \
@@ -635,7 +634,6 @@ def cf_variables(graphs, circ_singletons,out):
                 du.ETYPE_ID and '_%s' %i or ''),file=out)
         for v in G.nodes():
             print(' z{}_{}'.format(v, i),file=out)
-            print(' t{}'.format(v,i),file=out)
             for nm in ['n','m']:
                 for tp in SUBPATHTYPES:
                     print(' '+cf_var(mrd=i,v=v,mn=nm,tp=tp),file=out)

### 2023-10-21 scripts/spp_dcj.py
Revert "Fix some runtime exceptions; still untested though."
This reverts commit b87a386531e7c97c0bbd7cd98d10534c8c9b9441.

@@ -636,7 +636,7 @@ def cf_variables(graphs, circ_singletons,out):
             print(' z{}_{}'.format(v, i),file=out)
             for nm in ['n','m']:
                 for tp in SUBPATHTYPES:
-                    print(' '+cf_var(mrd=i,v=v,mn=nm,tp=tp),file=out)
+                    print(' '+cf_var(mrd=i,v=v,nm=nm,tp=tp),file=out)
             for tp in PATHTYPES:
                 print(' '+reportvar(mrd=i,v=v,tp=tp),file=out)
             for j in range(len(circ_singletons[(child, parent)])):

### 2023-10-21 scripts/data_utils.py
Revert "Fix runtime errors"
This reverts commit fb2c413385cf1075ac207750c744cdca3d05f805.

@@ -728,7 +728,7 @@ def _find_end_pairs(G, gName1, gName2):
     return {(u, v, Arun, Brun) for (u, v), (Arun, Brun) in res.items()}
 
 
-def checkGraph(G,capping=True):
+def checkGraph(G):
 
     for u, v, in G.edges():
         if u == v:
@@ -755,7 +755,7 @@ def checkGraph(G,capping=True):
             for data in G[u][v].values():
                 hasAdj |= data['type'] == ETYPE_ADJ
                 hasExtrOrId |= data['type'] in {ETYPE_ID, ETYPE_EXTR}
-        if not hasAdj and capping:
+        if not hasAdj:
             raise Exception(f'node {v} {G.nodes[v]["id"]} is not incident ' + \
                     'to an adjacency edge')
         if not hasExtrOrId:

### 2023-10-21 scripts/spp_dcj.py

@@ -96,7 +96,7 @@ def constraints(graphs, siblings, circ_singletons, caps, out):
 
     out.write('\n')
 
-def cf_constraints(graphs, siblings, circ_singletons, out):
+def cf_constraints(graphs, siblings, circ_singletons, caps, out):
 
     out.write('subject to\n')
 
@@ -119,7 +119,7 @@ def cf_constraints(graphs, siblings, circ_singletons, out):
         cfc13(G,i,out)
         cfc14_15(G,i,out)
         cfc16(G,i,out)
-        cfc17(circ_singletons,i,out)
+        cfc17(G,i,out)
     out.write('\n')
 
 
@@ -427,11 +427,8 @@ def cfc12(G,compnum,out):
 def cfc13(G,compnum,out):
     for v in G.nodes():
         for ij in PATHTYPES:
-            if ij == PTYPE_CYCLE:
-                #this constraint does not apply to cycles
-                continue
             for i,j in [tuple(ij),tuple(ij[::-1])]:
-                line = '{} + {} - {} <= 1\n'.format(mvar(v=v,mrd=compnum,tp=i),nvar(v=v,mrd=compnum,tp=j),reportvar(v=v,tp=ij,mrd=compnum))
+                line = '{} + {} - {} <= 1\n'.format(mvar(v=v,mrd=compnum,tp=i),nvar(v=v,mrd=compnum,tp=j))
                 out.write(line)
 
 def cfc14_15(G,compnum,out):
@@ -453,10 +450,10 @@ def cfc16(G,compnum,out):
     line = numerator+" - {}\n".format(summationvar(qr='q',mrd=compnum))
     out.write(line)
 
-def cfc17(circ_singletons, i, out):
+def cfc17(G,compnum,out,genomes):
     #for now handle circular singletons like in the original
     #TODO: Implement Barber shop variant
-    return c12(circ_singletons, i, out)
+    return c09(G, compnum, out, genomes)
 
 
 
@@ -472,7 +469,7 @@ def cf_objective(graphs,circ_singletons,alpha,beta,out):
     out.write(' + '.join(map(lambda x: '{} x{}'.format(x[1] * (1-alpha), \
             x[0]), adjs)))
     written=written or (len(adjs) > 0)
-    allnodeswithdata = dict([(v,data) for _,G in graphs.items() for v,data in G.nodes(data=True)])
+    allnodeswithdata = dict([(v,data) for _,G in graphs.items() for v,data in G.nodes()])
[…]

### 2023-10-21 scripts/data_utils.py
Revert "Stop ignoring caps when capping is on."
This reverts commit a9cadbb45e5183728e802fc769e8776840e843b1.

@@ -522,7 +522,7 @@ TWEIGHT = 'tweight'
 TPENALTY = 'tpenalty'
 
 def _constructRDAdjacencyEdges(G, gName, adjacencies, candidateWeights,
-        candidatePenalities, extremityIdManager,capping=True):
+        candidatePenalities, extremityIdManager):
     ''' create adjacencies of the genome named <gName>'''
     for ext1, ext2 in adjacencies:
 
@@ -530,13 +530,13 @@ def _constructRDAdjacencyEdges(G, gName, adjacencies, candidateWeights,
         penality = candidatePenalities.get((ext1, ext2), None)
         #Skip telomeres; we have not added them as vertices
         #but remember the weight of the telomeric adjacency
-        if is_telomere(ext1) and not capping:
+        if is_telomere(ext1):
             id2 = extremityIdManager.getId((gName, ext2))
             G.nodes[id2][TWEIGHT] = weight
             if penality is not None:
                 G.nodes[id2][TPENALTY]
             continue
-        elif is_telomere(ext2) and not capping:
+        elif is_telomere(ext2):
             id1 = extremityIdManager.getId((gName, ext1))
             G.nodes[id1][TWEIGHT] = weight
             if penality is not None:
@@ -906,7 +906,7 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
                 _constructRDTelomeres(G, gName, candidateTelomeres[gName],
                                     extremityIdManager)
             _constructRDAdjacencyEdges(G, gName, candidateAdjacencies[gName],
-                    candidateWeights, candidatePenalities, extremityIdManager,capping=capping)
+                    candidateWeights, candidatePenalities, extremityIdManager)
 
         fam2genes1 = mapFamiliesToGenes(genes[child], sep=sep)
         fam2genes2 = mapFamiliesToGenes(genes[parent], sep=sep)

### 2023-10-21 scripts/spp_dcj.py

@@ -732,7 +732,7 @@ if __name__ == '__main__':
             help='Separator of in gene names to split <family ID> and ' +
                     '<uniquifying identifier> in adjacencies file')
     
-    parser.add_argument('-cf','--capping-free',action='store_true',help='Activate (experimental) capping-free mode.',dest='cappingfree')
+    parser.add_argument('-cf','--capping-free',action='store_true',help='Activate (experimental) capping-free mode.',dest='cf')
 
     args = parser.parse_args()
 
@@ -741,6 +741,7 @@ if __name__ == '__main__':
     ch.setLevel(logging.INFO)
     ch.setFormatter(logging.Formatter('%(levelname)s\t%(asctime)s\t%(message)s'))
     LOG.addHandler(ch)
+    LOG.info("cf set to {}".format(args.cf))
     beta = args.beta
     if beta < 0:
         beta = args.alpha * 0.5
@@ -770,7 +771,7 @@ if __name__ == '__main__':
             'the tree').format(len(speciesTree)))
     relationalDiagrams = du.constructRelationalDiagrams(speciesTree,
             adjacencies, telomeres, weights, penalities, genes, ext2id,
-            sep=args.separator,capping=not args.cappingfree)
+            sep=args.separator,capping=not args.cf)
 
     graphs = relationalDiagrams['graphs']
 
@@ -818,31 +819,31 @@ if __name__ == '__main__':
         circ_singletons[ident] = du.identifyCircularSingletonCandidates(G)
         LOG.info(f'identified {len(circ_singletons[ident])} circular singleton candidates')
 
-    if not args.cappingfree:
+    if not args.cf:
         caps = getAllCaps(graphs)
     # construct & output ILP
     out = stdout
 
     LOG.info('writing objective over all graphs')
-    if args.cappingfree:
+    if args.cf:
         cf_objective(graphs,circ_singletons,args.alpha,beta,out)
     else:
         objective(graphs, circ_singletons, args.alpha, beta, out)
 
     LOG.info('writing constraints...')
-    if args.cappingfree:
+    if args.cf:
         cf_constraints(graphs, siblings, circ_singletons, out)
     else:
         constraints(graphs, siblings, circ_singletons, caps, out)
 
     LOG.info('writing domains...')
-    if args.cappingfree:
+    if args.cf:
         cf_domains(graphs,out)
     else:
         domains(graphs, out)
 
     LOG.info('writing variables...')
-    if args.cappingfree:
[…]

### 2023-10-21 scripts/data_utils.py
Revert "Fix bug"
This reverts commit 2db30a5f9de933df25e17d26dd80d1f785a363c4.

@@ -902,7 +902,7 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
 
         for gName in (child, parent):
             _constructRDNodes(G, gName, genes[gName], extremityIdManager)
-            if capping:
+            if not capping:
                 _constructRDTelomeres(G, gName, candidateTelomeres[gName],
                                     extremityIdManager)
             _constructRDAdjacencyEdges(G, gName, candidateAdjacencies[gName],

### 2023-10-21 scripts/spp_dcj.py

@@ -732,7 +732,7 @@ if __name__ == '__main__':
             help='Separator of in gene names to split <family ID> and ' +
                     '<uniquifying identifier> in adjacencies file')
     
-    parser.add_argument('-cf','--capping-free',action='store_true',help='Activate (experimental) capping-free mode.',dest='cf')
+    parser.add_argument('-cf','--capping-free',action='store_true',help='Activate (experimental) capping-free mode.')
 
     args = parser.parse_args()
 
@@ -741,7 +741,7 @@ if __name__ == '__main__':
     ch.setLevel(logging.INFO)
     ch.setFormatter(logging.Formatter('%(levelname)s\t%(asctime)s\t%(message)s'))
     LOG.addHandler(ch)
-    LOG.info("cf set to {}".format(args.cf))
+
     beta = args.beta
     if beta < 0:
         beta = args.alpha * 0.5

### 2023-10-21 scripts/data_utils.py
Revert "Implement basic capping-free ILP- completely untested"
This reverts commit 3e79dd3d1cc856e4286a0b44f343a0ea915c7017.

@@ -514,38 +514,17 @@ def mapFamiliesToGenes(genes, sep=DEFAULT_GENE_FAM_SEP):
     return res
 
 
-def is_telomere(r):
-    #TODO: Is this correct?
-    return r[1]=='o'
-
-TWEIGHT = 'tweight'
-TPENALTY = 'tpenalty'
-
 def _constructRDAdjacencyEdges(G, gName, adjacencies, candidateWeights,
         candidatePenalities, extremityIdManager):
     ''' create adjacencies of the genome named <gName>'''
     for ext1, ext2 in adjacencies:
-
-        weight = candidateWeights.get((ext1, ext2), 0)
-        penality = candidatePenalities.get((ext1, ext2), None)
-        #Skip telomeres; we have not added them as vertices
-        #but remember the weight of the telomeric adjacency
-        if is_telomere(ext1):
-            id2 = extremityIdManager.getId((gName, ext2))
-            G.nodes[id2][TWEIGHT] = weight
-            if penality is not None:
-                G.nodes[id2][TPENALTY]
-            continue
-        elif is_telomere(ext2):
-            id1 = extremityIdManager.getId((gName, ext1))
-            G.nodes[id1][TWEIGHT] = weight
-            if penality is not None:
-                G.nodes[id1][TPENALTY]
-            continue
         id1 = extremityIdManager.getId((gName, ext1))
         id2 = extremityIdManager.getId((gName, ext2))
+
         # ensure that each edge has a unique identifier
         edge_id = '{}_{}'.format(*sorted((id1, id2)))
+        weight = candidateWeights.get((ext1, ext2), 0)
+        penality = candidatePenalities.get((ext1, ext2), None)
         if penality is None:
             G.add_edge(id1, id2, type=ETYPE_ADJ, id=edge_id, weight=weight)
         else:
@@ -902,9 +881,8 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
 
         for gName in (child, parent):
             _constructRDNodes(G, gName, genes[gName], extremityIdManager)
-            if not capping:
-                _constructRDTelomeres(G, gName, candidateTelomeres[gName],
-                                    extremityIdManager)
+            _constructRDTelomeres(G, gName, candidateTelomeres[gName],
+                                  extremityIdManager)
             _constructRDAdjacencyEdges(G, gName, candidateAdjacencies[gName],
                     candidateWeights, candidatePenalities, extremityIdManager)
 

### 2023-10-21 scripts/spp_dcj.py

@@ -96,32 +96,6 @@ def constraints(graphs, siblings, circ_singletons, caps, out):
 
     out.write('\n')
 
-def cf_constraints(graphs, siblings, circ_singletons, caps, out):
-
-    out.write('subject to\n')
-
-    for i, ((child, parent), G) in enumerate(sorted(graphs.items())):
-
-        LOG.info(('writing capping-free constraints for relational diagram of {} and ' + \
-                '{}').format(child, parent))
-        cfc01(G,out)
-        cfc02(G,i,out)
-        cfc03(siblings[(child, parent)], out)
-        cfc04(G,i,out)
-        cfc05(G,i,out)
-        cfc06(G,i,out,(child,parent))
-        cfc07(G,i,out,(child,parent))
-        cfc08(G,i,out)
-        cfc09(G,i,out)
-        cfc10(G,i,out)
-        cfc11(G,i,out)
-        cfc12(G,i,out)
-        cfc13(G,i,out)
-        cfc14_15(G,i,out)
-        cfc16(G,i,out)
-        cfc17(G,i,out)
-    out.write('\n')
-
 
 def c01(G, out):
     for v, vdata in G.nodes(data = True):
@@ -455,57 +429,6 @@ def cfc17(G,compnum,out,genomes):
     #TODO: Implement Barber shop variant
     return c09(G, compnum, out, genomes)
 
-
-
-def cf_objective(graphs,circ_singletons,alpha,beta,out):
-    #TODO: What about telomeric, ie. artificial adjacencies?
-    #How to get their weight?
-    written=False
-    adjs = set(reduce(lambda x, y: x + y, (tuple(map(lambda z: (z[2]['id'], \
-            z[2]['weight']), filter(lambda x: x[2]['type'] == du.ETYPE_ADJ, \
-            G.edges(data = True)))) for i, (_, G) in \
-            enumerate(sorted(graphs.items())))))
-
-    out.write(' + '.join(map(lambda x: '{} x{}'.format(x[1] * (1-alpha), \
-            x[0]), adjs)))
-    written=written or (len(adjs) > 0)
-    allnodeswithdata = dict([(v,data) for _,G in graphs.items() for v,data in G.nodes()])
-    tlwght = ['{} t{}'.format((1-alpha)*data.get(du.TWEIGHT,0),v) for v,data in allnodeswithdata.items()]
-    if written and len(tlwght) > 0:
-        out.write(' + ')
-    out.write(' + '.join(tlwght))
-    written=written or (len(tlwght) > 0)
[…]

### 2023-10-21 scripts/data_utils.py
Revert "Implement all constraints but optimization constraints of cf ilp"
This reverts commit 3d573617a252fde090fd4733190db9a90539fb95.

@@ -870,7 +870,7 @@ def getIncidentAdjacencyEdges(G, v):
 
 def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
         candidateWeights, candidatePenalities, genes, extremityIdManager,
-        sep=DEFAULT_GENE_FAM_SEP,capping=True):
+        sep=DEFAULT_GENE_FAM_SEP):
     ''' constructs for each edge of the tree a relational diagram of the
     adjacent genomes'''
 
@@ -896,18 +896,17 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
 
 
     # create caps at last, assigning them the highest IDs
-    if capping:
-        for child, parent in tree:
-            G = res['graphs'][(child, parent)]
-            _constructRDCapping(G, child, parent, extremityIdManager)
-    #        _constructNaiveRDCapping(G, child, parent, extremityIdManager)
-            # remove caps from the graph that are not saturated by two edges
-    #        for v, d in tuple(G.degree()):
-    #            if d == 1:
-    #                if G.nodes[v]['type'] != VTYPE_CAP:
-    #                    raise Exception('assumption that all vertices with ' + \
-    #                            'degree one are caps failed')
-    #                G.remove_node(v)
+    for child, parent in tree:
+        G = res['graphs'][(child, parent)]
+        _constructRDCapping(G, child, parent, extremityIdManager)
+#        _constructNaiveRDCapping(G, child, parent, extremityIdManager)
+        # remove caps from the graph that are not saturated by two edges
+#        for v, d in tuple(G.degree()):
+#            if d == 1:
+#                if G.nodes[v]['type'] != VTYPE_CAP:
+#                    raise Exception('assumption that all vertices with ' + \
+#                            'degree one are caps failed')
+#                G.remove_node(v)
 
     return res
 

### 2023-10-21 scripts/spp_dcj.py

@@ -255,180 +255,6 @@ def c13(caps, out):
                 cap_set)), j), file = out)
 
 
-
-def cf_var(v=None,mrd=None,tp=None,mn=None):
-    if v==None or mrd==None or tp==None or mn==None:
-        #Variables are only optional so, I don't confuse their order.
-        raise Exception("Not enough values provided.")
-    return "{}v{}m{}t{}".format(mn,v,mrd,tp)
-
-def mvar(v=None,mrd=None,tp=None):
-    return cf_var(v=v,mrd=mrd,tp=tp,mn="m")
-
-def nvar(v=None,mrd=None,tp=None):
-    return cf_var(v=v,mrd=mrd,tp=tp,mn="n")
-
-def reportvar(v=None,mrd=None,tp=None):
-    return cf_var(v=v,mrd=mrd,tp=tp,mn="r")
-
-def summationvar(qr=None,mrd=None):
-    if qr==None or mrd==None:
-        #Variables are only optional so, I don't confuse their order.
-        raise Exception("Not enough values provided.")
-    return 'summationvar{}m{}'.format(qr,mrd)
-
-
-TELOMERE_A = 'A'
-TELOMERE_B = 'B'
-INDEL_A = 'a'
-INDEL_B = 'b'
-SUBPATHTYPES = [TELOMERE_A,TELOMERE_B,INDEL_A,INDEL_B]
-PTYPE_CYCLE = 'c'
-PATHTYPES = [PTYPE_CYCLE]+[a+b for a in SUBPATHTYPES for b in SUBPATHTYPES if a < b]
-
-def cfc01(G,out):
-    for v, vdata in G.nodes(data = True):
-        line = ' + '.join(["x{}".format(data['id']) for u in G.neighbors(v) 
-                           for data in G[u][v].values()
-                           if data['type']==du.ETYPE_ADJ ])
-        if line:
-            line+=" + "
-        line+="t{} = 1\n".format(v)
-        out.write(line)
-
-def cfc02(G, compnum, out):
-    for v in G.nodes():
-            #TODO: Why is there no comparison index necessary when the edge is an extremity edge?
-            #i.e. compare to:  line += 'x{}{}'.format(data['id'], data['type'] == du.ETYPE_ID and '_%s' %i or '')
-            line = ' + '.join(['x{}{}'.format(data['id'],'_%s'%compnum if data['type']==du.ETYPE_ID else '')
[…]

### 2023-10-21 scripts/data_utils.py
Revert "added parameter to set separator between gene family and uniquifying identifier"
This reverts commit b52325f02659933f51c30c8bff9e84e3348fd66d.

@@ -42,8 +42,6 @@ EXT_COMPLEMENT = {EXTR_HEAD:EXTR_TAIL,EXTR_TAIL:EXTR_HEAD}
 
 TRUE_ADJ_WEIGHT  = 1
 
-DEFAULT_GENE_FAM_SEP = '_'
-
 PAT_ADJ = re.compile('^(\w+)@([0-9_]+)$')
 PAT_MATCHED_EDGE = re.compile('^x(\d+)_(\d+)([^0-9 \t]*) 1\s*$')
 
@@ -91,12 +89,12 @@ def getTreeDepth(tree):
 
     return res
 
-def getFamiliesFromGenes(genesList,speciesList, sep=DEFAULT_GENE_FAM_SEP):
+def getFamiliesFromGenes(genesList,speciesList):
     resFamilies = {}
     for species in speciesList:
         resFamilies[species] = defaultdict(list)
         for gene in genesList[species]:
-            family = getFamily(gene, sep=sep)
+            family = getFamily(gene)
             resFamilies[species][family].append(gene)
     return(resFamilies)
 
@@ -112,7 +110,7 @@ def addAdjacency(ext1,ext2,weight,resAdjacencies,resWeights,resGenes):
     resGenes.append(gene2)
 
 
-def parseAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
+def parseAdjacencies(data):
     '''Read a file of adjacencies in the format species\tgene1\text1\tspecies\tgene2\text2\tweight'''
     headerMark = '#'
     delimiter = '\t'
@@ -142,14 +140,14 @@ def parseAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
     speciesList = list(resAdjacencies.keys())
     for species in speciesList:
         resGenes[species]    = list(set(resGenes[species]))
-    resFamilies = getFamiliesFromGenes(resGenes, speciesList, sep=sep)
+    resFamilies = getFamiliesFromGenes(resGenes,speciesList)
 
     return {'species':speciesList, 'genes':resGenes,
             'adjacencies':resAdjacencies, 'weights':resWeights,
             'families': resFamilies, 'penalities': resPenalities}
 
 
-def parseCandidateAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
+def parseCandidateAdjacencies(data):
     '''Read candidate adjacencies, returned as a dictionary, indexed by
     species, where for each species we have a list of pairs of gene
     extremities.
[…]

### 2023-10-21 scripts/spp_dcj.py

@@ -446,9 +446,6 @@ if __name__ == '__main__':
     parser.add_argument('-b', '--beta', default = -1, type=float,
             help='linear weighting factor for telomeric adjacencies;' + \
                     'if beta < 0, then beta is set to 1/2 * alpha')
-    parser.add_argument('-s', '--separator', default = du.DEFAULT_GENE_FAM_SEP, \
-            help='Separator of in gene names to split <family ID> and ' +
-                    '<uniquifying identifier> in adjacencies file')
 
     args = parser.parse_args()
 
@@ -466,11 +463,9 @@ if __name__ == '__main__':
     LOG.info('loading species tree from {}'.format(args.tree.name))
     speciesTree = du.parseTree(args.tree)
 
-    LOG.info(('loading candidate adjacencies from {}, using "{}" to separate' + \
-            ' gene family from uniquifying identifier').format(
-        args.candidateAdjacencies.name, args.separator))
-    candidateAdjacencies = du.parseAdjacencies(args.candidateAdjacencies,
-                                               sep=args.separator)
+    LOG.info('loading candidate adjacencies from {}'.format(
+        args.candidateAdjacencies.name))
+    candidateAdjacencies = du.parseAdjacencies(args.candidateAdjacencies)
 
     # add telomeres
     telomeres = identifyCandidateTelomeres(candidateAdjacencies,
@@ -486,8 +481,7 @@ if __name__ == '__main__':
     LOG.info(('constructing relational diagrams for all {} branches of ' + \
             'the tree').format(len(speciesTree)))
     relationalDiagrams = du.constructRelationalDiagrams(speciesTree,
-            adjacencies, telomeres, weights, penalities, genes, ext2id,
-            sep=args.separator)
+            adjacencies, telomeres, weights, penalities, genes, ext2id)
 
     graphs = relationalDiagrams['graphs']
 

### 2023-10-21 scripts/spp_dcj.py
start rough implementation

@@ -97,6 +97,9 @@ def constraints(graphs, siblings, circ_singletons, caps, out):
     out.write('\n')
 
 
+
+
+
 def c01(G, out):
     for v, vdata in G.nodes(data = True):
         line = ''
@@ -253,6 +256,21 @@ def c13(caps, out):
         if cap_set:
             print('{} - 2 a{} = 0'.format(' + '.join(map(lambda x: f'o{x}',
                 cap_set)), j), file = out)
+
+
+# implement constraints C.04 to C.36
+def mrd_constraints(graphs, siblings, circ_singletons, pseudocaps, out):
+
+    out.write('subject to\n')
+
+    for i, ((child, parent), G) in enumerate(sorted(graphs.items())):
+
+        LOG.info(('writing constraints for relational diagram of {} and ' + \
+                '{}').format(child, parent))
+        
+        
+
+    out.write('\n')
 
 
 def getAllCaps(graphs):

### 2023-10-21 docs/manuscript/main.tex
Use my default email address.

@@
 \institute{Faculty of Technology and Center for Biotechnology (CeBiTec), Bielefeld University, Germany\\ 
-    \email{lbohnenkaemper@techfak.uni-bielefeld.de}, \email{stoye@techfak.uni-bielefeld.de}\\
+    \email{lbohnenkaemper@techfak.uni-bielefeld.de}, \email{jens.stoye@uni-bielefeld.de}\\
 \and
@@
     \cns & $a_{\map(u)\map(v)} = x_{uv}$ & $\forall uv \in \adjacencyedges$ \\
-
+    \cns & $z_{v} \leq g_{\map (v)}$ & $\forall v \in \vertices$ \\
     (\ref{ilp:slmstart}) to (\ref{ilp:slmend})& Shao-Lin-Moret~\cite{SHA-LIN-MOR-2015} constraints& -- see Table~\ref{tab:slmcons}\\

### 2023-10-21 scripts/data_utils.py
Implement summary constraints. Add genome identification to data_utils. Missing still: constraints c01-c03,c22-c37, variables. Modifications to implement pseudo-capping with pseudocaps having the lowest ids and genome 1 having smaller ids than genome 2.

@@ -30,6 +30,9 @@ ETYPE_EXTR      = 'extremity'
 VTYPE_EXTR      = 'marker_extremity'
 VTYPE_CAP       = 'telomere'
 
+VGENOME_FIRST = 'A'
+VGENOME_SECOND = 'B'
+
 ORIENT_NEGATIVE = '-'
 ORIENT_POSITIVE = '+'
 
@@ -834,17 +837,17 @@ def _constructRDExtremityEdges(G, gName1, gName2, genes, fam2genes1,
     return siblings
 
 
-def _constructRDNodes(G, gName, genes, extremityIdManager):
+def _constructRDNodes(G, gName, genes, extremityIdManager,genome=VGENOME_FIRST):
     ''' create gene extremity nodes for the genome named <gName> '''
     for extr in (EXTR_HEAD, EXTR_TAIL):
         G.add_nodes_from(((extremityIdManager.getId((gName, (g, extr))),
-            dict(id=((gName, (g, extr))), type=VTYPE_EXTR)) for g in genes))
+            dict(id=((gName, (g, extr))), type=VTYPE_EXTR,genome=genome)) for g in genes))
 
 
-def _constructRDTelomeres(G, gName, telomeres, extremityIdManager):
+def _constructRDTelomeres(G, gName, telomeres, extremityIdManager,genome=VGENOME_FIRST):
     ''' create telomereic extremity nodes for the genome named <gName> '''
     G.add_nodes_from(((extremityIdManager.getId((gName, (t, 'o'))),
-        dict(id=((gName, (t, 'o'))), type=VTYPE_CAP)) for t in telomeres))
+        dict(id=((gName, (t, 'o'))), type=VTYPE_CAP,genome=genome)) for t in telomeres))
 
 
 def hasIncidentAdjacencyEdges(G, v):
@@ -877,9 +880,9 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
         G = nx.MultiGraph()
 
         for gName in (child, parent):
-            _constructRDNodes(G, gName, genes[gName], extremityIdManager)
+            _constructRDNodes(G, gName, genes[gName], extremityIdManager,genome=VGENOME_FIRST if gName==child else VGENOME_SECOND)
             _constructRDTelomeres(G, gName, candidateTelomeres[gName],
-                                  extremityIdManager)
+                                  extremityIdManager,genome=VGENOME_FIRST if gName==child else VGENOME_SECOND)
[…]

### 2023-10-21 scripts/spp_dcj.py

@@ -258,7 +258,23 @@ def c13(caps, out):
                 cap_set)), j), file = out)
 
 
-# implement constraints C.04 to C.36
+
+
+DEFAULT_SEP = '_'
+ENUMERATE_CONSTRAINTS = False
+
+#give a canonical string for a given edge between u,v with index k
+def disp_edge(u,v,k,sep=DEFAULT_SEP):
+    p = sep.join(sorted([u,v]))
+    return '{p}{s}{k}'.format(p=p,s=sep,k=k)
+
+
+
+'''
+I presume here that pseudo-caps in the graph are declared as du.VTYPE_CAP
+'''
+
+# implement constraints C.04 to C.37
 def mrd_constraints(graphs, siblings, circ_singletons, pseudocaps, out):
 
     out.write('subject to\n')
@@ -269,8 +285,78 @@ def mrd_constraints(graphs, siblings, circ_singletons, pseudocaps, out):
                 '{}').format(child, parent))
         
         
+        
 
     out.write('\n')
+
+
+def cfc04(i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
+    print("{enm}f{s}{e} - n{s}{e} + c{s}{e} - q{s}{e} - s{s}{e} = 0".format(s=sep,e=i,enm="c04: " if enm else ""),file=out)
+
+def cfc05(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
+    sumstr = ' + '.join(['0.5 x{s}{i}{s}{e}'.format(s=sep,e=disp_edge(u,v,k),i=i) for u,v,k,tp in G.edges(keys=True,data=type) if tp==du.ETYPE_EXTR])
+    print("{enm}{sumstr} - n{s}{e} = 0".format(sumstr=sumstr,e=i,s=sep,enm="c05: " if enm else ""),file=out)
+
+def cfc06(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
+    sumstr = ' + '.join(['rc{s}{i}{s}{v}'.format(i=i,v=v,s=sep) for v,vt in G.nodes(data='type') if vt != du.VTYPE_CAP])
[…]

### 2023-10-21 scripts/data_utils.py
Some changes to implement SLM constraints - need to check with the others first.

@@ -786,7 +786,7 @@ def rotateToMin(path):
 
 
 def _constructRDExtremityEdges(G, gName1, gName2, genes, fam2genes1,
-        fam2genes2, extremityIdManager):
+        fam2genes2, extremityIdManager,triple_sibs_format=False):
 
     genes1 = genes[gName1]
     genes2 = genes[gName2]
@@ -814,11 +814,14 @@ def _constructRDExtremityEdges(G, gName1, gName2, genes, fam2genes1,
             edge_idh = '{}_{}'.format(*sorted((id1h, id2h)))
             edge_idt = '{}_{}'.format(*sorted((id1t, id2t)))
 
-            G.add_edge(id1h, id2h, type=ETYPE_EXTR, id=edge_idh)
-            G.add_edge(id1t, id2t, type=ETYPE_EXTR, id=edge_idt)
+            k1 = G.add_edge(id1h, id2h, type=ETYPE_EXTR, id=edge_idh)
+            k2 = G.add_edge(id1t, id2t, type=ETYPE_EXTR, id=edge_idt)
 
             # ensure sorted order of sibling edges
-            siblings.append((edge_idh, edge_idt))
+            if not triple_sibs_format:
+                siblings.append((edge_idh, edge_idt))
+            else:
+                siblings.append(((id1h, id2h,k1),(id1t, id2t,k2)))
 
         # create indel edges between genes of smaller family
         for i, gName in enumerate((gName1, gName2)):
@@ -870,7 +873,7 @@ def getIncidentAdjacencyEdges(G, v):
 
 
 def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
-        candidateWeights, candidatePenalities, genes, extremityIdManager):
+        candidateWeights, candidatePenalities, genes, extremityIdManager,triple_sibs_format=False):
     ''' constructs for each edge of the tree a relational diagram of the
     adjacent genomes'''
 
@@ -889,7 +892,7 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
         fam2genes1 = mapFamiliesToGenes(genes[child])
         fam2genes2 = mapFamiliesToGenes(genes[parent])
         siblings   = _constructRDExtremityEdges(G, child, parent, genes,
-                fam2genes1, fam2genes2, extremityIdManager)
+                fam2genes1, fam2genes2, extremityIdManager,triple_sibs_format=triple_sibs_format)
 
[…]

### 2023-10-21 scripts/spp_dcj.py

@@ -355,7 +355,29 @@ def cfc20(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
 
 def cfc21(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
     for cc, v in enumerate(G.nodes(),start=1):
-        print("z{s}{i}{s}{v} - g{s}{v} <= 0".format(i=i,v=v,s=sep,enm="c.21.%d: "%cc if enm else ""),file=out)
+        print("{enm}z{s}{i}{s}{v} - g{s}{v} <= 0".format(i=i,v=v,s=sep,enm="c.21.%d: "%cc if enm else ""),file=out)
+
+
+def cfc22(sibs,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
+    for cc,es in enumerate(sibs):
+        e1 = disp_edge(*es[0])
+        e2 = disp_edge(*es[1])
+        print("{enm}x{s}{i}{s}{e1} - x{s}{i}{s}{e2} = 0".format(i=i,e1=e1,e2=e2,s=sep,enm="c22.%d: "%cc if enm else ""),file=out)
+
+'''TODO: implement this after checking w/ dany
+def cfc23(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
+    for cc,pacde in enumerate(G.edges(keys=True,data=True)):
+        u_,v_,k,data = pacde
+        e = disp_edge(u_,v_,k)
+        for u,v in [(u_,v_),(v_,u_)]:
+            if data['type'] != du.ETYPE_ID:
+                print("{enm}y{s}{i}{s}{u} - y{s}{i}{s}{v} + {ixu} x{s}{i}{s}{e}  <= {ixu}".format(
+                    i=i,u=u,v=v,e=e,ixu=G.nodes[u]['id']
+                ))
+def cfc24(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
+    pass
+'''
+
 
 
 

### 2023-10-21 docs/manuscript/abstract.tex
Extract some sections into their own files.

--- 2023-10-20 docs/manuscript/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.21 of 2022/01/12
-%
-\documentclass[runningheads]{llncs}
-%
-\input{preamble}
-\usepackage[T1]{fontenc}
-% T1 fonts will be used to generate the final print and online PDFs,
-% so please use T1 fonts in your manuscript whenever possible.
-% Other font encondings may result in incorrect characters.
-%
-\usepackage{graphicx}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following two lines
-% to display URLs in blue roman font according to Springer's eBook style:
-
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%
-
-\begin{document}
-%
-\title{Reconstructing Rearrangement Phylogenies of Natural Genomes}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{Leonard Bohnen\"amper\inst{1}\orcidID{0000-0003-4508-0078} \and
-    Jens Stoye \inst{1}\orcidID{0000-0002-4656-7155} \and
-    Daniel D\"orr \inst{2}\orcidID{0000-0002-3720-6227}
-}
-%
-\authorrunning{L. Bohnenk\"amper et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Faculty of Technology and Center for Biotechnology (CeBiTec), Bielefeld University, Germany\\ 
-    \email{lbohnenkaemper@techfak.uni-bielefeld.de}, \email{stoye@techfak.uni-bielefeld.de}\\
-\and
-Institute for Medical Biometry and Bioinformatics, Medical Faculty, and Center for Digital Medicine, Heinrich Heine University, Germany\\
-\email{daniel.doerr@hhu.de}}
-
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
     We study the classic problem of inferring ancestral genomes on a given phylogeny and a given set of extant genomes, also known as \emph{small parsimony problem} (SPP). 
@@
     Even under simple evolutionary models, this problem is computationally intractable,
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-\section{Introduction}
-\section{Background}
-\subsection{Problem Formulation}
-\section{A new Method}
-\subsection{Pre-selecting Adjacencies}
-\todoDany{Describe Process}
-\subsection{A New ILP Formulation}
-\todoLeo{ILP goes here}
-\todoLeo{Remove idx stuff and bring back $v_i$ notation}
-
-\begin{algorithm}
-\caption{Capping-free Small Parsimony}
-\textbf{Objective}
-
-\newcommand{\idx}{\texttt{ix}}
-\hspace{0.5cm}\texttt{Maximize}
-\begin{equation*}
-    \alpha \sum_{\edt\in\tree} f_\edt + .... \text{\todoDany{Optimization of other factors goes here.}}
-\end{equation*}
-
-\textbf{For each ancestral genome $\X \in \tree$ with $\ggraph(\X) = (\vertices, \adjacencyedges)$}
-
-\begin{constraints}
-\begin{tabular}{lcl}
-    \cns & $\minfam{F} \leq \sum_{v \in F} g_v \leq \maxfam{F}$ & $\forall F \in \families$\\
-    \cns & $g_v = g_u$ & for each pair of nodes $(u, v)$ corresponding to head/tail extremity of the same gene\todoDany{improve}\\
-%    \cns & $g_{v} = x^{\X\A}_{v} = x^{\X\B}_{v} = x^{\X\C}_{v}$& $\forall v \in V$ and genomes $\A$, $\B$, $\C$ adjacent to $\X$ in $\Gamma$\\
-    \cns & $\sum_{uv \in \adjacencyedges} a_{uv} = g_u$ & $\forall u \in \vertices$\\
-\end{tabular}
-\end{constraints}
-
-\medskip
-\textbf{For each tree edge $(\A,\B):=\edt\in \tree$ with $\pcmrd(\A,\B) = (\vertices\cup\pseudocaps, \adjacencyedges \cup \extremityedges \cup \selfedges)$:}
-
-\begin{constraints}
-\begin{tabular}{lcl}
-
-	\cns & $f_\edt = n_\edt - c_\edt + q_\edt + s_\edt$\\
-	\cns & $n_\edt = 0.5 \sum_{e \in \extremityedges} x_{e}$\\
-    \cns & $c_\edt = \sum_{v\in \vertices} r^c_{v}$\\
-    \cns & $2q_\edt \geq p^{a\trans b}+ p^{ABa} + p^{ABb} - p^{A\trans B} $\\
-    \cns & $p^{a\trans b} = \sum_{v \in \vertices}r^{a\trans b}_{v}$\\
-    \cns & $p^{A\trans b} = \sum_{v \in \pseudocaps^A} r^{A\trans b}_e$\\
-    \cns & $p^{B\trans a} = \sum_{v\in \pseudocaps^B} r^{B\trans a}_e$\\
-    \cns & $p^{A\cis a} = \sum_{v \in \pseudocaps^A} r^{A\cis a}_e$\\
-    \cns & $p^{B\cis b} = \sum_{e \in \pseudocaps^B} r^{B\cis b}_e$\\
-    \cns & $p^{ABa} \geq p^{B\trans a}$\\
-    \cns & $p^{ABa} \geq p^{A\cis a}$\\
-    \cns & $p^{ABb} \geq p^{B\cis b}$\\
-    \cns & $p^{ABb} \geq p^{A\trans b}$\\
-    \cns & $p^{A\trans B} = \sum_{v\in \pseudocaps^A} r^{A\trans B}_v$\\
-    \cns & $s_\edt = \sum_{v\in\vertices} r^{s}_{v}$ \\
-    \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= g_{\map(u)}$ & $\forall u\in \vertices$  \\
-    \cns & $a_{\map(u)\map(v)} = x_{uv}$ & $\forall uv \in \adjacencyedges$ \\
-
-    (\ref{ilp:slmstart}) to (\ref{ilp:slmend})& Shao-Lin-Moret~\cite{SHA-LIN-MOR-2015} constraints& -- see Table~\ref{tab:slmcons}\\
-    (\ref{ilp:regvstart}) to (\ref{ilp:regvend})& Reporting for regular vertices& -- see Table~\ref{tab:regv}\\
-    (\ref{ilp:pcstart}) to (\ref{ilp:pcend})& Reporting for \gls{pcaps}& -- see Table~\ref{tab:pcaps}\\
-    (\ref{ilp:csstart}) to (\ref{ilp:csend})& Reporting circular singletons& --  see Table~\ref{tab:csreport}\\    
-\end{tabular}
-
-\end{constraints}
-\todoLeo{Put domains here?}
-\todoLeo{Fix formatting}
-\end{algorithm}
-
-\newcommand{\idx}{}
-\begin{table}
-
-\begin{constraints}
-\caption{Shao-Lin-Moret constraints.} \label{tab:slmcons}
-\begin{tabular}{lcl}
-% \cns\label{ilp:slmstart} & $\sum_{uv\in \ades} x_{e} = 1 $ & $\forall u\in \vertices$\\
-    \cns\label{ilp:slmstart} & $x_e=x_d$ & for all sibling edges $e,d$\\
-    \cns & $y_v + \idx{u}(1-x_{uv}) \geq y_u$ &$\forall uv \in \adjacencyedges\cup \extremityedges$\\
-         & $\idx{u}(1-x_{uv})\geq y_u$& $\forall uv \in \selfedges$\\
-    \cns\label{ilp:slmend} & $\idx{v}z_v \leq y_v$ & $\forall v\in \vertices\cup\pseudocaps $\\
-\end{tabular}
-\end{constraints}
-
-\end{table}
-
-\begin{table}
-
-\begin{constraints}
-\caption{Reporting for regular vertices.\comment{This is basically exactly the same as in ding, just swap out $r^{ab}$ for $t$ and $\parityvar$ for $r$ and root reports on vertices instead of edges.}} \label{tab:regv}
-\begin{tabular}{lcl}
-\cns\label{ilp:regvstart} & $\parityvar^\edt_v \leq 1 - x_{uv}$ & $\forall uv \in \selfedges^\A$\\
-     & $\parityvar^\edt_v \geq  x_{uv}$ & $\forall uv \in \selfedges^\B$\\
-\cns & $\parityvar_v \leq \parityvar_u +  (1-x_{uv})$& $\forall uv \in \extremityedges$\\
-& $\parityvar_u \leq \parityvar_v + r_{uv}^{a\trans b} + (1-x_{uv})$& $\forall uv\in E_{adj}^\A,u,v\notin \pseudocaps$\\
-& $\parityvar_u \leq \parityvar_v + (1-x_{uv})$& $\forall uv\in E_{adj}^\B,u,v\notin \pseudocaps$\\
-
-\cns & $r_{v}^c \leq z_v$&$\forall v \in \vertices^\A$\\
-\cns\label{ilp:regvend} & $r_{u}^{a\trans b} \leq x_{uv}$&$\forall uv\in\selfedges^\A$\\
-\end{tabular}
-\end{constraints}
-
-\end{table}
-
-\begin{table}
-
-\begin{constraints}
-\caption{Reporting for \gls{pcaps}.} \label{tab:pcaps}
-\begin{tabular}{lcl}
-\cns\label{ilp:pcstart} & $\parityvar_v = 0$ & $\forall v \in \pseudocaps^\A$\\
- & $\parityvar_v = 1$ & $\forall v \in \pseudocaps^\B$\\
-\cns & $\parityvar_u \leq \parityvar_v + r_{v}^{A\trans B} + r_{v}^{A\trans b} + (1-x_{uv})$& $\forall uv\in \ades, v \in \pseudocaps^\A$\\
-    & $\parityvar_u \leq \parityvar_v + r_{u}^{B\trans a} + (1-x_{uv})$& $\forall uv\in \ades, u\in \pseudocaps^\B$\\
-    \cns & $r_{v}^{A\trans B} \leq z_v$&$\forall v \in \pseudocaps^\A$\\
-    \cns & $1 -y_v \leq r^{A\trans b}_{v} + r^{A\cis a}_v$ & $v \in \pseudocaps^\A$\\
-        & $1 -y_v \leq r^{B\trans a}_{v} + r^{B\cis b}_v$ & $v \in \pseudocaps^\B$\\
-    \cns & $y_{v_i} \leq i(1-r_{v}^{R})$&$v\in \pseudocaps^\A,R\in  \{A\trans b, A\cis a\}$\\
-     & $y_{v_i} \leq i(1-r_{v}^{R})$&$v\in \pseudocaps^\B,R\in  \{B\trans a, B\cis b\}$\\
-    \cns\label{ilp:pcend} & $r_{v}^{A\trans B} \leq b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^A$ \\
-    & $r_{v}^{A\trans b} \leq b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^A$ \\
-    & $r_{v}^{B\trans a} \leq 1-b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^\B$\\
-\end{tabular}
-\end{constraints}
-
-\end{table}
-
-\begin{table} \caption{Reporting circular singletons.\comment{(This is what we developed last week.)}}\label{tab:csreport}
-\begin{tabular}{lcl}
-     \cns\label{ilp:csstart} & $d_u+d_v + x_{uv}\leq 2$ & $\forall uv \in \adjacencyedges\cup \selfedges$  \\
-     & $d_u + d_v - x_{uv} \geq 0$ & $\forall uv \in \adjacencyedges\cup \selfedges$  \\
-    \cns & $w_u = w_v$ & $\forall uv \in \selfedges$\\
-    \cns\label{ilp:csend} &$K (1- x_{uv} + r^{s}_{u} + r^{s}_{v}) + w_v \geq w_u + d_v - d_u $ & $\forall uv\in \adjacencyedges$\\
-     %& $K (1- x_{uv} + r^{s}_{v} + r^{s}_{u}) + w_u \geq w_v + d_u - d_v $ & $\forall uv\in \adjacencyedges$\\
-\end{tabular}
-\end{table}
-
-\todoLeo{Fix inconsistencies with comparison in $\parityvar$ and $z$.}
-
-\section{Evaluation}
-\section{Discussion}
-
-\subsubsection{Acknowledgements} Please place your acknowledgments at
-the end of the paper, preceded by an unnumbered run-in heading (i.e.
-3rd-level heading).
-
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-\bibliographystyle{splncs04}
-\bibliography{refs}
-
-\end{document}

### 2023-10-21 docs/manuscript/main.tex
Extract some sections into their own files.

@@
 \begin{abstract}
-    We study the classic problem of inferring ancestral genomes on a given phylogeny and a given set of extant genomes, also known as \emph{small parsimony problem} (SPP). 
-    More specifically, we address a scenario where genomes represent sets of linear or circular successions of oriented genomic markers orders, also known as \emph{gene orders} that evolve along the branches of the phylogeny and are altered by a general evolutionary model that supports a genome rearrangements emulated through the double-cut-and-join (DCJ) operation and by segmental insertions and deletions.
 
-    Even under simple evolutionary models, this problem is computationally intractable,
+\input{abstract}
 
@@
 \subsection{A New ILP Formulation}
-\todoLeo{ILP goes here}
-\todoLeo{Remove idx stuff and bring back $v_i$ notation}
 
-\begin{algorithm}
-\caption{Capping-free Small Parsimony}
-\textbf{Objective}
-
-\newcommand{\idx}{\texttt{ix}}
-\hspace{0.5cm}\texttt{Maximize}
-\begin{equation*}
-    \alpha \sum_{\edt\in\tree} f_\edt + .... \text{\todoDany{Optimization of other factors goes here.}}
-\end{equation*}
-
-\textbf{For each ancestral genome $\X \in \tree$ with $\ggraph(\X) = (\vertices, \adjacencyedges)$}
-
-\begin{constraints}
-\begin{tabular}{lcl}
-    \cns & $\minfam{F} \leq \sum_{v \in F} g_v \leq \maxfam{F}$ & $\forall F \in \families$\\
-    \cns & $g_v = g_u$ & for each pair of nodes $(u, v)$ corresponding to head/tail extremity of the same gene\todoDany{improve}\\
-%    \cns & $g_{v} = x^{\X\A}_{v} = x^{\X\B}_{v} = x^{\X\C}_{v}$& $\forall v \in V$ and genomes $\A$, $\B$, $\C$ adjacent to $\X$ in $\Gamma$\\
-    \cns & $\sum_{uv \in \adjacencyedges} a_{uv} = g_u$ & $\forall u \in \vertices$\\
-\end{tabular}
-\end{constraints}
-
-\medskip
-\textbf{For each tree edge $(\A,\B):=\edt\in \tree$ with $\pcmrd(\A,\B) = (\vertices\cup\pseudocaps, \adjacencyedges \cup \extremityedges \cup \selfedges)$:}
-
-\begin{constraints}
-\begin{tabular}{lcl}
-
-	\cns & $f_\edt = n_\edt - c_\edt + q_\edt + s_\edt$\\
-	\cns & $n_\edt = 0.5 \sum_{e \in \extremityedges} x_{e}$\\
-    \cns & $c_\edt = \sum_{v\in \vertices} r^c_{v}$\\
-    \cns & $2q_\edt \geq p^{a\trans b}+ p^{ABa} + p^{ABb} - p^{A\trans B} $\\
-    \cns & $p^{a\trans b} = \sum_{v \in \vertices}r^{a\trans b}_{v}$\\
-    \cns & $p^{A\trans b} = \sum_{v \in \pseudocaps^A} r^{A\trans b}_e$\\
-    \cns & $p^{B\trans a} = \sum_{v\in \pseudocaps^B} r^{B\trans a}_e$\\
-    \cns & $p^{A\cis a} = \sum_{v \in \pseudocaps^A} r^{A\cis a}_e$\\
-    \cns & $p^{B\cis b} = \sum_{e \in \pseudocaps^B} r^{B\cis b}_e$\\
-    \cns & $p^{ABa} \geq p^{B\trans a}$\\
-    \cns & $p^{ABa} \geq p^{A\cis a}$\\
-    \cns & $p^{ABb} \geq p^{B\cis b}$\\
-    \cns & $p^{ABb} \geq p^{A\trans b}$\\
-    \cns & $p^{A\trans B} = \sum_{v\in \pseudocaps^A} r^{A\trans B}_v$\\
-    \cns & $s_\edt = \sum_{v\in\vertices} r^{s}_{v}$ \\
-    \cns & $\sum_{uv\in E_{ext} \cup E_{id} } x_{uv}= g_{\map(u)}$ & $\forall u\in \vertices$  \\
-    \cns & $a_{\map(u)\map(v)} = x_{uv}$ & $\forall uv \in \adjacencyedges$ \\
-    \cns & $z_{v} \leq g_{\map (v)}$ & $\forall v \in \vertices$ \\
-    (\ref{ilp:slmstart}) to (\ref{ilp:slmend})& Shao-Lin-Moret~\cite{SHA-LIN-MOR-2015} constraints& -- see Table~\ref{tab:slmcons}\\
-    (\ref{ilp:regvstart}) to (\ref{ilp:regvend})& Reporting for regular vertices& -- see Table~\ref{tab:regv}\\
-    (\ref{ilp:pcstart}) to (\ref{ilp:pcend})& Reporting for \gls{pcaps}& -- see Table~\ref{tab:pcaps}\\
-    (\ref{ilp:csstart}) to (\ref{ilp:csend})& Reporting circular singletons& --  see Table~\ref{tab:csreport}\\    
-\end{tabular}
-
-\end{constraints}
-\todoLeo{Put domains here?}
-\todoLeo{Fix formatting}
-\end{algorithm}
-
-\newcommand{\idx}{}
-\begin{table}
-
-\begin{constraints}
-\caption{Shao-Lin-Moret constraints.} \label{tab:slmcons}
-\begin{tabular}{lcl}
-% \cns\label{ilp:slmstart} & $\sum_{uv\in \ades} x_{e} = 1 $ & $\forall u\in \vertices$\\
-    \cns\label{ilp:slmstart} & $x_e=x_d$ & for all sibling edges $e,d$\\
-    \cns & $y_v + \idx{u}(1-x_{uv}) \geq y_u$ &$\forall uv \in \adjacencyedges\cup \extremityedges$\\
-         & $\idx{u}(1-x_{uv})\geq y_u$& $\forall uv \in \selfedges$\\
-    \cns\label{ilp:slmend} & $\idx{v}z_v \leq y_v$ & $\forall v\in \vertices\cup\pseudocaps $\\
-\end{tabular}
-\end{constraints}
-
-\end{table}
-
-\begin{table}
-
-\begin{constraints}
-\caption{Reporting for regular vertices.\comment{This is basically exactly the same as in ding, just swap out $r^{ab}$ for $t$ and $\parityvar$ for $r$ and root reports on vertices instead of edges.}} \label{tab:regv}
-\begin{tabular}{lcl}
-\cns\label{ilp:regvstart} & $\parityvar^\edt_v \leq 1 - x_{uv}$ & $\forall uv \in \selfedges^\A$\\
-     & $\parityvar^\edt_v \geq  x_{uv}$ & $\forall uv \in \selfedges^\B$\\
-\cns & $\parityvar_v \leq \parityvar_u +  (1-x_{uv})$& $\forall uv \in \extremityedges$\\
-& $\parityvar_u \leq \parityvar_v + r_{uv}^{a\trans b} + (1-x_{uv})$& $\forall uv\in E_{adj}^\A,u,v\notin \pseudocaps$\\
-& $\parityvar_u \leq \parityvar_v + (1-x_{uv})$& $\forall uv\in E_{adj}^\B,u,v\notin \pseudocaps$\\
-
-\cns & $r_{v}^c \leq z_v$&$\forall v \in \vertices^\A$\\
-\cns\label{ilp:regvend} & $r_{u}^{a\trans b} \leq x_{uv}$&$\forall uv\in\selfedges^\A$\\
-\end{tabular}
-\end{constraints}
-
-\end{table}
-
-\begin{table}
-
-\begin{constraints}
-\caption{Reporting for \gls{pcaps}.} \label{tab:pcaps}
-\begin{tabular}{lcl}
-\cns\label{ilp:pcstart} & $\parityvar_v = 0$ & $\forall v \in \pseudocaps^\A$\\
- & $\parityvar_v = 1$ & $\forall v \in \pseudocaps^\B$\\
-\cns & $\parityvar_u \leq \parityvar_v + r_{v}^{A\trans B} + r_{v}^{A\trans b} + (1-x_{uv})$& $\forall uv\in \ades, v \in \pseudocaps^\A$\\
-    & $\parityvar_u \leq \parityvar_v + r_{u}^{B\trans a} + (1-x_{uv})$& $\forall uv\in \ades, u\in \pseudocaps^\B$\\
-    \cns & $r_{v}^{A\trans B} \leq z_v$&$\forall v \in \pseudocaps^\A$\\
-    \cns & $1 -y_v \leq r^{A\trans b}_{v} + r^{A\cis a}_v$ & $v \in \pseudocaps^\A$\\
-        & $1 -y_v \leq r^{B\trans a}_{v} + r^{B\cis b}_v$ & $v \in \pseudocaps^\B$\\
-    \cns & $y_{v_i} \leq i(1-r_{v}^{R})$&$v\in \pseudocaps^\A,R\in  \{A\trans b, A\cis a\}$\\
-     & $y_{v_i} \leq i(1-r_{v}^{R})$&$v\in \pseudocaps^\B,R\in  \{B\trans a, B\cis b\}$\\
-    \cns\label{ilp:pcend} & $r_{v}^{A\trans B} \leq b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^A$ \\
-    & $r_{v}^{A\trans b} \leq b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^A$ \\
-    & $r_{v}^{B\trans a} \leq 1-b_u$&$\forall uv\in \adjacencyedges, v \in \pseudocaps^\B$\\
-\end{tabular}
-\end{constraints}
-
-\end{table}
-
-\begin{table} \caption{Reporting circular singletons.\comment{(This is what we developed last week.)}}\label{tab:csreport}
-\begin{tabular}{lcl}
-     \cns\label{ilp:csstart} & $d_u+d_v + x_{uv}\leq 2$ & $\forall uv \in \adjacencyedges\cup \selfedges$  \\
-     & $d_u + d_v - x_{uv} \geq 0$ & $\forall uv \in \adjacencyedges\cup \selfedges$  \\
-    \cns & $w_u = w_v$ & $\forall uv \in \selfedges$\\
-    \cns\label{ilp:csend} &$K (1- x_{uv} + r^{s}_{u} + r^{s}_{v}) + w_v \geq w_u + d_v - d_u $ & $\forall uv\in \adjacencyedges$\\
-     %& $K (1- x_{uv} + r^{s}_{v} + r^{s}_{u}) + w_u \geq w_v + d_u - d_v $ & $\forall uv\in \adjacencyedges$\\
-\end{tabular}
-\end{table}
-
-\todoLeo{Fix inconsistencies with comparison in $\parityvar$ and $z$.}
-
+\include{ilp}
 \section{Evaluation}

### 2023-10-21 docs/manuscript/abstract.tex
Rewrite abstract.

@@
-    We study the classic problem of inferring ancestral genomes on a given phylogeny and a given set of extant genomes, also known as \emph{small parsimony problem} (SPP). 
-    More specifically, we address a scenario where genomes represent sets of linear or circular successions of oriented genomic markers orders, also known as \emph{gene orders} that evolve along the branches of the phylogeny and are altered by a general evolutionary model that supports a genome rearrangements emulated through the double-cut-and-join (DCJ) operation and by segmental insertions and deletions.
+    We study the classic problem of inferring ancestral genomes on a given phylogeny and a given set of extant genomes, also known as the \emph{small parsimony problem} (SPP).
+    More specifically, we address a scenario that considers the evolving structure of a genome via large scale rearrangements.
+    To abstract from local mutations, such as SNPs, genomes are represented by their \emph{gene orders} on multiple linear or circular chromosomes.
+    Hereby, we do not limit ourselves to genomes with a particular gene set, nor do we restrict the multiplicity of each gene per genome.
+    Since gene orders and multiplicities may differ among extant genomes, the aim of the SPP in this context is to find marker orders and multiplicities for ancestral genomes which are most parsimonious under a given model.
+    The model we consider is a general evolutionary model that supports genome rearrangements emulated through the double-cut-and-join~(DCJ) operation and by segmental insertions and deletions.
 
-    Even under simple evolutionary models, this problem is computationally intractable,
+Even under simple evolutionary models, such as the classic character-state model, the SPP is computationally intractable.
+However, we give a highly optimized ILP that is able to compute the SPP under our model for sufficiently small phylogenies.

### 2023-10-22 docs/manuscript/abstract.tex
rewritten abstract

--- 2023-10-21 docs/manuscript/abstract.tex
@@
-    We study the classic problem of inferring ancestral genomes on a given phylogeny and a given set of extant genomes, also known as \emph{small parsimony problem} (SPP). 
-    More specifically, we address a scenario where genomes represent sets of linear or circular successions of oriented genomic markers orders, also known as \emph{gene orders} that evolve along the branches of the phylogeny and are altered by a general evolutionary model that supports a genome rearrangements emulated through the double-cut-and-join (DCJ) operation and by segmental insertions and deletions.
+    We study the classic problem of inferring ancestral genomes under a given phylogeny and from a given set of extant genomes, also known as the \emph{small parsimony problem} (SPP).
+    The evolutionary model considered at hand encompasses large scale rearrangements acting on so-called \emph{gene order} sequences and includes segmental gain and loss.
+Each genome may encompass one or more linear or circular chromosomes, and genes may appear in several copies, without restriction on their genomic location or orientation within the sequence.
+    Therefore, our method presumes ancestral copy number estimates, specified either as fixed count, or as range. 
+    In case of the latter, our method chooses the copy number within the range that minimizes the total number of large scale rarrangements over the entire phylogeny.
 
-    Even under simple evolutionary models, this problem is computationally intractable,
+Even under simple evolutionary models, such as the classic character-state model, the SPP is computationally intractable.
+However, we give a highly optimized ILP that is able to compute the SPP under our model for sufficiently small phylogenies and gene families.
+We benchmark our method on simulated phylogenies and discuss its performance in reconstructing gene oders under our broad evolutionary model.

### 2023-10-23 scripts/data_utils.py
Implement local constraints

@@ -840,17 +840,21 @@ def _constructRDExtremityEdges(G, gName1, gName2, genes, fam2genes1,
     return siblings
 
 
-def _constructRDNodes(G, gName, genes, extremityIdManager,genome=VGENOME_FIRST):
+def _constructRDNodes(G, gName, genes, extremityIdManager,genome=VGENOME_FIRST,localIdManager=None):
     ''' create gene extremity nodes for the genome named <gName> '''
+    if localIdManager is None:
+        localIdManager = extremityIdManager
     for extr in (EXTR_HEAD, EXTR_TAIL):
         G.add_nodes_from(((extremityIdManager.getId((gName, (g, extr))),
-            dict(id=((gName, (g, extr))), type=VTYPE_EXTR,genome=genome)) for g in genes))
+            dict(id=((gName, (g, extr))), type=VTYPE_EXTR,genome=genome,localid=localIdManager.getId((gName, (g, extr))))) for g in genes))
 
 
-def _constructRDTelomeres(G, gName, telomeres, extremityIdManager,genome=VGENOME_FIRST):
+def _constructRDTelomeres(G, gName, telomeres, extremityIdManager,genome=VGENOME_FIRST,localIdManager=None):
     ''' create telomereic extremity nodes for the genome named <gName> '''
+    if localIdManager is None:
+        localIdManager = extremityIdManager
     G.add_nodes_from(((extremityIdManager.getId((gName, (t, 'o'))),
-        dict(id=((gName, (t, 'o'))), type=VTYPE_CAP,genome=genome)) for t in telomeres))
+        dict(id=((gName, (t, 'o'))), type=VTYPE_CAP,genome=genome,localid=localIdManager.getId((gName, (t,'o'))))) for t in telomeres))
 
 
 def hasIncidentAdjacencyEdges(G, v):

### 2023-10-23 scripts/spp_dcj.py

@@ -364,22 +364,151 @@ def cfc22(sibs,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
         e2 = disp_edge(*es[1])
         print("{enm}x{s}{i}{s}{e1} - x{s}{i}{s}{e2} = 0".format(i=i,e1=e1,e2=e2,s=sep,enm="c22.%d: "%cc if enm else ""),file=out)
 
-'''TODO: implement this after checking w/ dany
+#TODO: implement this after checking w/ dany
 def cfc23(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
-    for cc,pacde in enumerate(G.edges(keys=True,data=True)):
+    cc = 1
+    for pacde in G.edges(keys=True,data=True):
         u_,v_,k,data = pacde
         e = disp_edge(u_,v_,k)
         for u,v in [(u_,v_),(v_,u_)]:
             if data['type'] != du.ETYPE_ID:
                 print("{enm}y{s}{i}{s}{u} - y{s}{i}{s}{v} + {ixu} x{s}{i}{s}{e}  <= {ixu}".format(
-                    i=i,u=u,v=v,e=e,ixu=G.nodes[u]['id']
-                ))
+                    i=i,u=u,v=v,e=e,ixu=G.nodes[u]['localid'],s=sep,enm='c23.%d '%cc if enm else ""
+                ),file=out)
+                cc+=1
+            else:
+                print("{enm}y{s}{i}{s}{u} + {ixu} x{s}{i}{s}{e}  <= {ixu}".format(
+                    i=i,u=u,e=e,ixu=G.nodes[u]['localid'],s=sep,enm='c23.%d '%cc if enm else ""
+                ),file=out)
+                cc+=1
+
 def cfc24(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
-    pass
-'''
+    for cc,pacdv in enumerate(G.nodes(data='localid')):
+        v,lid = pacdv
+        print("{enm}{lid} z{s}{i}{s}{v} - y{s}{i}{s}{v} <= 0".format(i=i,v=v,lid=lid,s=sep,enm="c24.%d: "%cc if enm else ''),file=out)
 
+def cfc25(G,i,out,sep=DEFAULT_SEP,enm=ENUMERATE_CONSTRAINTS):
+    cc = 1
[…]

### 2023-10-23 docs/manuscript/abstract.tex
Fix some typos in abstract and ILP.

@@
     Therefore, our method presumes ancestral copy number estimates, specified either as fixed count, or as range. 
-    In case of the latter, our method chooses the copy number within the range that minimizes the total number of large scale rarrangements over the entire phylogeny.
+    In case of the latter, our method chooses the copy number within the range that minimizes the total number of large scale rearrangements over the entire phylogeny.
 
@@
 However, we give a highly optimized ILP that is able to compute the SPP under our model for sufficiently small phylogenies and gene families.
-We benchmark our method on simulated phylogenies and discuss its performance in reconstructing gene oders under our broad evolutionary model.
+We benchmark our method on simulated phylogenies and discuss its performance in reconstructing gene orders under our broad evolutionary model.

### 2023-10-24 docs/manuscript/main.tex
intermediate save

@@
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{genome rearrangement \and ancestral reconstruction \and small parsimony \and integer linear programming \and double-cut-and-join}
 \end{abstract}
-%
-%
-%
 
-\section{Introduction}
+%\section{Introduction}
 \section{Background}
-\subsection{Problem Formulation}
+
+\begin{itemize}
+    \item SPP algoirthms such as 
+\end{itemize}
 \section{A new Method}
+\subsection{Preliminaries}
+General notation
+\begin{itemize}
+    \item Weighted multigraph $G = (\vertices, \edges, \w)$ with edge weight function $\w: \edges \to \mathbb R$
+    \item A \emph{phylogeny} $\tree$ is a connected graph. Nodes of $\tree$ with degree 1 are termed \emph{leaves}.
+\end{itemize}
+
+Paper-specific notation:
+
+\begin{itemize}
+    \item Genomic marker 
+    \item Genome $\X$
+    \item Genome adjacency graph $\ggraph(\X) = (V, \adjacencyedges)$ \todoDany{find better name?}
+    \item 
+We model family assignments of marker extremities as a function $\family: \extremities \to \mathbb N$ for which holds true that for any marker $g = \{g^\et, g^\eh\}$, $\family(g^\et) = \family(g^\eh)$.
+Function 
+$\minmulti_\X\varepsilon)$ reports the minimum multiplicity of the marker corresponding to extremity $\varepsilon$ in a given genome $\X$, while $\maxmulti_\X(\varepsilon)$ reports its maximum multiplicity.
+A family assignment $\family'$ is \emph{$\family$-derived} if there exists no two extremities $\varepsilon, \varepsilon'$ with $\family'(\varepsilon) = \family'(\varepsilon')$ and $\family(\varepsilon) \neq \family(\varepsilon')$.
+A family assignment $\family$ is termed \emph{$\mathcal \A$-resolved} (or simply \emph{resolved} if the context is clear) if for each member $\X$ of a set of genomes $\mathcal A$, each extremity $\varepsilon \in \extremities $ has a multiplicity not higher than $1$, \emph{i.e.}, $\maxmulti_\X(\varepsilon)\leq 1$.
+
+    \item Multirelational diagram $\mrd$ and \emph{capping-free multi-relational diagram} (CFMRD) $\pcmrd$
+\end{itemize}
 \subsection{Pre-selecting Adjacencies}
@@
 \subsection{A New ILP Formulation}
+The algorithm described in the following operates on two levels: 
+on the local level, a genome is derived from each genome adjacency graph, constituting a set of linear or circular chromosomes. 
+On the global level, genomes are connected to each other along the branches of the phylogeny. Each branch gives rise to a pairwise comparison by means of the CFMRD. 
+In doing so, the selection of adjacencies of a derived genome is propagated from across CFMRDs, thus ensuring global consistency.
+
+\paragraph{Local level.} Constraint \ref{c:cn}
+
+\paragraph{Global level.}
+When comparing pairs of genomes within a CFMRD, we make use of a capping-free formulation for the computation of the pairwise DCJ indel distance~\cite{BOH-2023}.
+
+\todoLeo{Insert a description of the formulation here}
 
 \include{ilp}
+
 \section{Evaluation}

### 2023-10-24 docs/manuscript/abstract.tex
intermediate save

@@
     The evolutionary model considered at hand encompasses large scale rearrangements acting on so-called \emph{gene order} sequences and includes segmental gain and loss.
-Each genome may encompass one or more linear or circular chromosomes, and genes may appear in several copies, without restriction on their genomic location or orientation within the sequence.
+Each genome may encompass one or more linear or circular chromosomes, and genomic markers may appear in several copies, without restriction on their genomic location or orientation within the sequence.
     Therefore, our method presumes ancestral copy number estimates, specified either as fixed count, or as range. 

### 2023-10-25 docs/manuscript/main.tex
changed typeface of ilp variables

@@
 Function 
-$\minmulti_\X\varepsilon)$ reports the minimum multiplicity of the marker corresponding to extremity $\varepsilon$ in a given genome $\X$, while $\maxmulti_\X(\varepsilon)$ reports its maximum multiplicity.
-A family assignment $\family'$ is \emph{$\family$-derived} if there exists no two extremities $\varepsilon, \varepsilon'$ with $\family'(\varepsilon) = \family'(\varepsilon')$ and $\family(\varepsilon) \neq \family(\varepsilon')$.
-A family assignment $\family$ is termed \emph{$\mathcal \A$-resolved} (or simply \emph{resolved} if the context is clear) if for each member $\X$ of a set of genomes $\mathcal A$, each extremity $\varepsilon \in \extremities $ has a multiplicity not higher than $1$, \emph{i.e.}, $\maxmulti_\X(\varepsilon)\leq 1$.
-
+$\minmulti_\X : \mathbb N \to \mathbb N$ reports the minimum multiplicity of a gene family in a given genome $\X$, while $\maxmulti_\X: \mathbb N \to \mathbb N$ reports its maximum multiplicity. 
     \item Multirelational diagram $\mrd$ and \emph{capping-free multi-relational diagram} (CFMRD) $\pcmrd$
@@
 
-\paragraph{Local level.} Constraint \ref{c:cn}
+\paragraph{Local level.} Constraint \ref{c:cn} implements the user-provided copy number constraints for each gene family $i$, where $F^A_i = \{ v \mid v \in V: v \text{ belongs to }\A\text{ and } \family(v) = i\}$ and $F^B_i$ defined analogously.
 

### 2023-10-25 docs/manuscript/main.tex
worked a little bit on describing the local level of the ILP

@@
     \item Genomic marker 
+    \item Furthermore, we use a function $\extf : \extremities \to \{\et, \eh, \circ\}$ to map extremities to their corresponding kind (tail, head or telomere). 
     \item Genome $\X$
@@
 
-\paragraph{Local level.} Constraint \ref{c:cn} implements the user-provided copy number constraints for each gene family $i$, where $F^A_i = \{ v \mid v \in V: v \text{ belongs to }\A\text{ and } \family(v) = i\}$ and $F^B_i$ defined analogously.
+\paragraph{Local level.} 
+Constraint \ref{c:cn} implements the user-provided copy number requirement. 
+In doing so, we define the set of tail extremities of genes from family $i$ in the GAG of genome $\X$ as $F_i = \{ v \mid v \in V: \family(v) = i \text{ and } \extf(v) = \et\}$ and iterate over the associated $\ilpvar{g}$ variables of its members to bound the number of active genes within the given range.
+The choice of counting tail extremities is arbitrary, however, the subsequent constraint (\ref{c:cn_consistent}) ensures that the assignment of $\ilpvar{g}$ variables for both extremities of a gene is consistent.
+At last, Constraint \ref{c:genome} makes sure that a valid genome is derived from the GAG when solving the ILP.
 

### 2023-10-25 docs/manuscript/main.tex
use package for todos

--- 2023-10-25 docs/manuscript/main.tex
@@
     \item Genomic marker 
+    \item Furthermore, we use a function $\extf : \extremities \to \{\et, \eh, \circ\}$ to map extremities to their corresponding kind (tail, head or telomere). 
     \item Genome $\X$
-    \item Genome adjacency graph $\ggraph(\X) = (V, \adjacencyedges)$ \todoDany{find better name?}
+    \item Genome adjacency graph $\ggraph(\X) = (\vertices, \edges, \family)$ \todoD{find better name?}
     \item 
@@
 \subsection{Pre-selecting Adjacencies}
-\todoDany{Describe Process}
+\todoD{Describe Process}
 \subsection{A New ILP Formulation}
 The algorithm described in the following operates on two levels: 
-on the local level, a genome is derived from each genome adjacency graph, constituting a set of linear or circular chromosomes. 
+on the local level, a genome is derived from each GAG, constituting a set of linear or circular chromosomes. 
 On the global level, genomes are connected to each other along the branches of the phylogeny. Each branch gives rise to a pairwise comparison by means of the CFMRD. 
@@
 
-\paragraph{Local level.} Constraint \ref{c:cn} implements the user-provided copy number constraints for each gene family $i$, where $F^A_i = \{ v \mid v \in V: v \text{ belongs to }\A\text{ and } \family(v) = i\}$ and $F^B_i$ defined analogously.
+\paragraph{Local level.} 
+Constraint \ref{c:cn} implements the user-provided copy number requirement. 
+In doing so, we define the set of tail extremities of markers from family $i$ in GAG $\ggraph(\X)$ as $F_i = \{ v \mid v \in V: \family(v) = i \text{ and } \extf(v) = \et\}$ and iterate over the associated $\ilpvar{g}$ variables of its members to bound the number of active markers within the given range.
+The choice of counting tail extremities is arbitrary, however, the subsequent constraint (\ref{c:cn_consistent}) ensures that the assignment of $\ilpvar{g}$ variables for both extremities of a markers is consistent.
+At last, Constraint \ref{c:genome} makes sure that a valid genome is derived from the GAG when solving the ILP, by enforcing that each extremity node of $\ggraph(\X)$ is incident to at most one adjacency edge.
 
 \paragraph{Global level.}
+
 When comparing pairs of genomes within a CFMRD, we make use of a capping-free formulation for the computation of the pairwise DCJ indel distance~\cite{BOH-2023}.
 
-\todoLeo{Insert a description of the formulation here}
+\todoL{Insert a description of the formulation here}
 

### 2023-11-03 docs/manuscript/main.tex
improved notation; we are getthing there!

@@
 
-%\section{Introduction}
+\section{Introduction}
 \section{Background}
 
+\subsection{Preliminaries}
+A \emph{(genomic) marker} $g := (g^\et, g^\eh)$ is a universally unique entity consisting of \emph{marker extremities} tail of $g$, denoted by $g^\et$, and head of $g$, denoted by $g^\eh$. 
+A \emph{telomere} $t^\circ$ is a universally unique entity encompassing a single telomeric extremity denoted by ``$\circ$''. 
+A genome is a triple $(M \cup \pseudocaps, A, \family)$, with $M$ being its marker set and $\pseudocaps$ its set of telomeres, $A \subset (M \cup \pseudocaps)^2$ the set of telomeres, and $\family: M \to \mathbb N$ a function indicating the family of each marker.
+The set of adjacencies $A$ has the properties that \emph{(i)} for each extremity $\varepsilon \in \bigcup M \cup \pseudocaps$, there exists exactly one adjacency $a \in A$ that contains $\varepsilon$.
+The adjacency set decomposes the set of markers and telomeres into linear or circular components, that we further call \emph{chromosomes}. 
+Moreover, \emph{(ii)} each linear chromosome, formed by ordering markers and telomeres based on their adjacencies in $A$, starts and ends with a telomere.
+
+A \emph{(genome) scaffold} is a quintuple $(M \cup \pseudocaps, A, \family, \multi, \w)$ with 
 \begin{itemize}
-    \item SPP algoirthms such as 
+    \item $M$, $\pseudocaps$, and $\family$ as defined in a genome, 
+    \item the set of adjacencies $A$ satisfying that for each $g^\et \in \bigcup A$, there exists also extremity $g^\eh \in \bigcup A$ and vice versa, and each telomeric extremity is used only once, i.e., $\forall \{X, X'\} \subseteq A$, $X \cap X' \cap \pseudocaps = \emptyset$,
+    \item copy number function $\multi(i) = [a, b]$ reporting for each family $i$ the permitted range $0 \leq a \leq b$ of its copy number, such that each family, defined as $F_i := \{g \in M \mid f(g) = i\}$, satisfies $|F_i| = b$, and
+    \item adjacency weight function $\w: A \to \mathbb R$. % [-1, 1]$.
 \end{itemize}
-\section{A new Method}
-\subsection{Preliminaries}
+
+Observe that a genome can give rise to a scaffold, with copy number $\multi$ and weight $\w$ functions arbitrarily defined, but the reverse does not hold true in general.
+% not sure we need this:
+% However, in the following, we regard genomes as scaffolds with tight copy numbers, that is, the copy number range of each family $i$ corresponds to a point that coincides with the number of its associated markers.
+A genome $S' = (M', A', \family')$ is \emph{derived} from a scaffold $S = (M, A, \family, \multi, \w)$, or simply ``$S$-derived'', if \emph{(i)} $M' \subseteq M$, \emph{(ii)} $A' \subseteq A$, \emph{(iii)}
+there exist no two markers $g, g' \in M'$ with $\family'(g) = \family'(g')$ and $\family(g) \neq \family(g')$, and \emph{(iv)} the original copy number constraints are satisfied, i.e., for each family $i$ and the set of markers associated with $i$ in scaffold $S$, i.e., $F_i = \{ g \in M' \mid \family(g) = i\}$, holds true that $|F_i| \in \multi(i)$.
+We call a scaffold $S$ \emph{linearizable} if there exists an $S$-derived genome. 
+In fact, many scaffolds are not linearizable.
+
+A \emph{phylogeny} $\tree$ is a connected graph with nodes representing \emph{operational taxonomic units} (OTUs). 
+Nodes with degree 1 are termed \emph{tips} of the phylogeny.
+
 General notation
 \begin{itemize}
+    \item 
     \item Weighted multigraph $G = (\vertices, \edges, \w)$ with edge weight function $\w: \edges \to \mathbb R$
-    \item A \emph{phylogeny} $\tree$ is a connected graph. Nodes of $\tree$ with degree 1 are termed \emph{leaves}.
 \end{itemize}
@@
     \item Furthermore, we use a function $\extf : \extremities \to \{\et, \eh, \circ\}$ to map extremities to their corresponding kind (tail, head or telomere). 
-    \item Genome $\X$
-    \item Genome adjacency graph $\ggraph(\X) = (\vertices, \edges, \family)$ \todoD{find better name?}
     \item 
@@
 \end{itemize}
+
+\begin{itemize}
+    \item SPP algoirthms such as 
+\end{itemize}
+
+\begin{problem}[Weighted CN-constrained degenerate DCJ indel distance]\label{prb:wdeg_dcj}
+    Given a weighting scheme $\w : \extremities \times \extremities \to \mathbb R$, some $\alpha \in [0, 1]$, two linearizable degenerate genomes $\A, \B$ with copy number constraints \hl{XX}\todoD{add data structure} and family assignment $\family$, find $\A$-derived genome $\A'$, $\B$-derived genome $\B'$, and $\family$-derived $\{\A',\B'\}$-resolved family assignment $\family'$ that minimize the linear combination
+    $$
+    (1-\alpha) \cdot \sum_{X \in \A' \cup \B'} -\w(X) + \alpha \cdot \dist_\DCJid(A', B')\,.
+    $$
+\end{problem}
+
+\begin{problem}[SPP-DCJ]\label{prb:spp_dcj}
+    Given a phylogeny $\tree$ and a set of linearizable degenerate genomes $\A_1, \ldots, \A_k$ corresponding to the node set $V(\tree) = \{\A_1, \ldots, \A_k\}$, find genomes $\A'_1 \subseteq \A_1, \ldots, \A'_k \allowbreak \subseteq \A_k$ that minimize the sum of weighted degenerate DCJ indel distances along the edges of $\tree$.
+\end{problem}
+
+\section{A new Method}
 \subsection{Pre-selecting Adjacencies}
-\todoD{Describe Process}
-\subsection{A New ILP Formulation}
+\todoD{Describe Process}\subsection{A New ILP Formulation}
+
 The algorithm described in the following operates on two levels: 

### 2023-12-04 docs/manuscript/main.tex
Revert "improved notation; we are getthing there!"
This reverts commit 5847de2da673d15a31f58b0f2476cb11dc976303.

--- 2023-10-25 docs/manuscript/main.tex
@@
 
-%\section{Introduction}
+\section{Introduction}
 \section{Background}
 
+\subsection{Preliminaries}
+A \emph{phylogeny} $\tree$ is a connected graph with nodes representing \emph{operational taxonomic units} (OTUs). 
+Nodes with degree 1 are termed \emph{tips} of the phylogeny.
+A \emph{genome scaffold} of an OTU $\A$ is a quintuple $(M \cup \pseudocaps, A, \family, \multi, \w)$ with 
 \begin{itemize}
-    \item SPP algoirthms such as 
+    \item $M$ being the marker set of $\A$ and $\pseudocaps$ its set of telomeres,
+    \item $A$ its set of unique adjacencies satisfying 
+        \emph{(i)} $\forall g^\et \in \bigcup A$, there exists also extremity $g^\eh \in \bigcup A$ and vice versa, and 
+        \emph{(ii)} each telomeric extremity is used only once: $\forall \{X, X'\} \subseteq A$, $X \cap X' \cap \pseudocaps = \emptyset$. 
+    \item $\family: M \to \mathbb N$ a function indicating the gene family of each marker of $\A$,
+    \item a function $\multi(i) = [a, b]$ that reports for each family $i$ the permitted range $0 \leq a \leq b$ of its copy number, such that each family $F_i = \{m \in M \mid f(m) = i\}$ satisfies $|F_i| = b$,
+    \item and adjacency weight function $\w: A \to \mathbb R$.
 \end{itemize}
-\section{A new Method}
-\subsection{Preliminaries}
+% \begin{enumerate}
+% \end{enumerate}
+A genome is a triple $(M \cup \pseudocaps, A, \family)$ for which the following holds true:
+\begin{itemize}
+    \item The extremities of each marker $m = (m^\et, m^\eh) \in M$ occur in exactly one adjacency, i.e., $|\{ a \in A \mid m^\et \in a\}| = |\{ a \in A \mid m^\eh \in a\}| = 1$,
+    \item 
+\end{itemize}
+Note that a genome is also a scaffold, but the reverse does not hold true in general.
+
+A genome $A'$ is \emph{$A$-derived} from degenerate genome $A$ (or simply ``derived from $A$'') if $A' \subseteq A$ and $\bigcup A' \setminus \pseudocaps = \bigcup A \setminus \pseudocaps$. 
+Conversely, a degenerate genome $A$ is \emph{linearizable} if there exists an $A$-derived genome. 
+In fact, many degenerate genomes are not linearizable.
+
 General notation
 \begin{itemize}
+    \item 
     \item Weighted multigraph $G = (\vertices, \edges, \w)$ with edge weight function $\w: \edges \to \mathbb R$
-    \item A \emph{phylogeny} $\tree$ is a connected graph. Nodes of $\tree$ with degree 1 are termed \emph{leaves}.
 \end{itemize}
@@
     \item Genome $\X$
-    \item Genome adjacency graph $\ggraph(\X) = (\vertices, \edges, \family)$ \todoD{find better name?}
+    \item \hl{Genome adjacency graph}\todoD{replace with degenerate genome} $\ggraph(\X) = (\vertices, \edges, \family)$ 
     \item 

### 2023-12-05 scripts/displayEventCounts.py
update script to newer pandas version

@@ -91,7 +91,7 @@ def readEventCounts(events_folder, rev_tree_dict):
 
     branch_dict['Overall'] = total
     for branch in branch_dict:
-        event_counts = event_counts.append({ \
+        event_counts = pd.concat([event_counts, pd.DataFrame(data={ \
                 'Branch':           branch, \
                 'Dup_events':       branch_dict[branch][0], \
                 'Dup_genes':        branch_dict[branch][1], \
@@ -99,7 +99,7 @@ def readEventCounts(events_folder, rev_tree_dict):
                 'Lost_genes':       branch_dict[branch][3], \
                 'Inversions':       branch_dict[branch][4], \
                 'Transpositions':   branch_dict[branch][5] \
-                },ignore_index=True)
+                }, index=isinstance(branch, tuple) and range(len(branch)) or [0])],ignore_index=True)
 
     return event_counts
 

### 2023-12-05 scripts/noisy_adjacencies.py
latest version of noisy_adjacencies

@@ -2,960 +2,155 @@
 
 # import from built-in packages
 from argparse import ArgumentParser, ArgumentDefaultsHelpFormatter as ADHF, FileType
-from collections import defaultdict, deque
-from itertools import product, chain, combinations
-from sys import stderr
+from sys import stdout, stderr, exit
+from collections import defaultdict
 import random
 import csv
-import re
 
+# import from own packages
+import data_utils as du
 
-# import from third-party packages
-import networkx as nx
+NOISY_ADJ_WEIGHT = 1
+ADVERSARIAL_ADJ_WEIGHT = 1
 
-#
-# global variables
-#
+MAX_TRIES = 10000 # Max number of tries to find a random adjacency
 
-CHR_CIRCULAR    = ')'
-CHR_LINEAR      = '|'
-EXTR_HEAD       = 'h'
-EXTR_TAIL       = 't'
-
-ETYPE_ADJ       = 'adj'
-ETYPE_ID        = 'id'
-ETYPE_EXTR      = 'extremity'
-
-VTYPE_EXTR      = 'marker_extremity'
-VTYPE_CAP       = 'telomere'
-
-ORIENT_NEGATIVE = '-'
-ORIENT_POSITIVE = '+'
-
-SIGN2EXT_1      = {ORIENT_NEGATIVE:EXTR_TAIL,ORIENT_POSITIVE:EXTR_HEAD}
-SIGN2EXT_2      = {ORIENT_NEGATIVE:EXTR_HEAD,ORIENT_POSITIVE:EXTR_TAIL}
-EXT2SIGN_1      = {EXTR_TAIL:ORIENT_NEGATIVE,EXTR_HEAD:ORIENT_POSITIVE}
-EXT2SIGN_2      = {EXTR_TAIL:ORIENT_POSITIVE,EXTR_HEAD:ORIENT_NEGATIVE}
-
-EXT_COMPLEMENT = {EXTR_HEAD:EXTR_TAIL,EXTR_TAIL:EXTR_HEAD}
-
-TRUE_ADJ_WEIGHT  = 1
-
-PAT_ADJ = re.compile('^(\w+)@([0-9_]+)$')
-PAT_MATCHED_EDGE = re.compile('^x(\d+)_(\d+)([^0-9 \t]*) 1\s*$')
-
-
-#
-# DATA ACQUISITION, PARSERS
-#
-
-def parseTree(data):
-    ''' Reads a tree returned as a list of branches [descendant parent] '''
[…]

### 2023-12-05 scripts/spp_dcj.py
remove penality stuff

@@ -59,11 +59,11 @@ def objective(graphs, circ_singletons, alpha, beta, out):
         # subtract circular singleton penality
         out.write(' - '.join(('{} s{}_{}'.format(alpha, j, i) for j in range(len(cs)))))
 
-    for adj, weight in set(reduce(lambda x, y: x + y, (tuple(map(lambda z: (z[2]['id'], \
-            z[2]['penality']), filter(lambda x: 'penality' in x[2] and \
-            x[2]['type'] == du.ETYPE_ADJ, G.edges(data = True)))) for i, (_, G) in \
-            enumerate(sorted(graphs.items()))))):
-        out.write(f' - {weight * beta} x{adj}')
+#    for adj, weight in set(reduce(lambda x, y: x + y, (tuple(map(lambda z: (z[2]['id'], \
+#            z[2]['penality']), filter(lambda x: 'penality' in x[2] and \
+#            x[2]['type'] == du.ETYPE_ADJ, G.edges(data = True)))) for i, (_, G) in \
+#            enumerate(sorted(graphs.items()))))):
+#        out.write(f' - {weight * beta} x{adj}')
 
     out.write('\n\n')
 
@@ -480,13 +480,12 @@ if __name__ == '__main__':
     genes = candidateAdjacencies['genes']
     adjacencies = candidateAdjacencies['adjacencies']
     weights = candidateAdjacencies['weights']
-    penalities = candidateAdjacencies['penalities']
 
     ext2id = du.IdManager()
     LOG.info(('constructing relational diagrams for all {} branches of ' + \
             'the tree').format(len(speciesTree)))
     relationalDiagrams = du.constructRelationalDiagrams(speciesTree,
-            adjacencies, telomeres, weights, penalities, genes, ext2id,
+            adjacencies, telomeres, weights, genes, ext2id,
             sep=args.separator)
 
     graphs = relationalDiagrams['graphs']

### 2024-05-11 scripts/data_utils.py
Implement testing and fix ILP bugs known so far

@@ -319,6 +319,7 @@ def unimog2adjacencies(genome):
 
     # ignore genome name
     _, chromosomes = genome
+    tels=0
     for chr_ in chromosomes:
         # set counter for first marker
         if chr_[1][0][1] not in occ:
@@ -326,7 +327,6 @@ def unimog2adjacencies(genome):
         occ[chr_[1][0][1]] += 1
 
         fst_occ = occ[chr_[1][0][1]]
-
         for i in range(len(chr_[1])-1):
             (o1, g1), (o2, g2) = chr_[1][i:i+2]
             if g2 not in occ:
@@ -341,8 +341,10 @@ def unimog2adjacencies(genome):
             res.append(((f'{g1}_{occ[g1]}', SIGN2EXT_1[o1]),
                 (f'{g2}_{fst_occ}', SIGN2EXT_2[o2])))
         elif chr_[0] == CHR_LINEAR:
-            res.append(((f'{g1}_{occ[g1]}', SIGN2EXT_1[o1]), ('t', EXTR_HEAD)))
-            res.append((('t', EXTR_HEAD), (f'{g2}_{fst_occ}', SIGN2EXT_2[o2])))
+            tels+=1
+            res.append(((f'{g1}_{occ[g1]}', SIGN2EXT_1[o1]), ('t_{}'.format(tels), EXTR_CAP)))
+            tels+=1
+            res.append((('t_{}'.format(tels), EXTR_CAP), (f'{g2}_{fst_occ}', SIGN2EXT_2[o2])))
 
     return res
 
@@ -710,7 +712,8 @@ def _find_end_pairs(G, gName1, gName2):
 
 
 def checkGraph(G,cf=False):
-
+    for v,data in G.nodes(data=True):
+        print(v,",".join(["{}={}".format(k,x) for k,x in data.items()]),file=sys.stderr)
     for u, v, in G.edges():
         if u == v:
             raise Exception(f'node {v} is connected to itself')
@@ -839,12 +842,13 @@ def _constructRDExtremityEdges(G, gName1, gName2, genes, fam2genes1,
 
     return siblings
 
-
+TELOMERE_GENE='t'
 def _constructRDNodes(G, gName, genes, globalIdManager,localIdManager):
     ''' create gene extremity nodes for the genome named <gName> '''
+    
     for extr in (EXTR_HEAD, EXTR_TAIL):
         G.add_nodes_from(((localIdManager.getId((gName, (g, extr))),
[…]

### 2024-05-11 scripts/spp_dcj.py

@@ -52,7 +52,7 @@ def constraints(graphs, siblings, out):
                 '{}').format(child, parent))
         genomes = [child,parent]
         global_constraints(G,out)
-        loose_constraints(G,i,out)
+        loose_constraints(G,i,genomes,out)
         slm_constraints(G,i,siblings,out)
         regular_reporting(G,i,genomes,out)
         pcap_reporting(G,i,genomes,out)
@@ -84,8 +84,6 @@ def c02(G,out):
 
 def c03(G,out):
     for v,data in G.nodes(data=True):
-        if data['type']==du.VTYPE_CAP:
-            continue
         aes = [G[v][u][i]['anc'] for u in G[v] for i in G[v][u] if G[v][u][i]['type']==du.ETYPE_ADJ]
         sm = ' + '.join(['a{sep}{e}'.format(sep=du.SEP,e=e) for e in aes])
         print("{sm} - g{sep}{v} = 0".format(sm=sm,sep=du.SEP,v=G.nodes[v]['anc']),file=out)
@@ -107,29 +105,29 @@ def c06(G,tree_edge,out):
     xs = ['0.5 x{sep}{te}{sep}{e}'.format(te=tree_edge,sep=du.SEP,e=data['id']) for u,v,data in G.edges(data=True) if data['type']==du.ETYPE_EXTR]
     print('{s} - n{sep}{te} = 0'.format(s=' + '.join(xs),sep=du.SEP,te=tree_edge),file=out)
 
-def c07(G,tree_edge,out):
-    cs = ['rc{sep}{te}{sep}{v}'.format(te=tree_edge,sep=du.SEP,v=v) for v,data in G.nodes(data=True) if data['type'] != du.VTYPE_CAP]
+def c07(G,tree_edge,genomes,out):
+    cs = ['rc{sep}{te}{sep}{v}'.format(te=tree_edge,sep=du.SEP,v=v) for v,data in G.nodes(data=True) if data['type'] != du.VTYPE_CAP and get_genome(G,v)==genomes[0]]
     print('{s} - c{sep}{te} = 0'.format(s=' + '.join(cs),sep=du.SEP,te=tree_edge),file=out)
 
 def c08(G,tree_edge,out):
[…]

### 2024-05-12 scripts/sol_tester.py
Further test cases.

@@ -1,0 +1,45 @@
+#!/usr/bin/env python3
+
+from argparse import ArgumentParser
+import sys
+parser = ArgumentParser()
+parser.add_argument("solfile")
+parser.add_argument("--solval",type=float)
+parser.add_argument("--vars",type=str,nargs='*',default=[])
+parser.add_argument("--vals",type=int,nargs='*',default=[])
+
+args = parser.parse_args()
+
+
+
+valdict = dict(zip(args.vars,args.vals))
+
+ok =True
+with open(args.solfile) as f:
+    obj_txt = '# Objective value = '
+    for line in f:
+        if line.startswith(obj_txt):
+            obj_value = float(line[len(obj_txt):])
+            if args.solval:
+                if abs(args.solval-obj_value) > 0.001:
+                    ok=False
+                    print("Wrong objective value, expected {} but got {}.".format(args.solval,obj_value))
+            continue
+        elif line.startswith('#'):
+            continue
+        v,val_ = line.strip().split()
+        #print(v,val_)
+        val=int(val_)
+        if v not in valdict:
+            continue
+        if valdict[v]!=val:
+            ok=False
+            print("Wrong value, expexted {}={}, but got {}.".format(v,valdict[v],val))
+        del valdict[v]
+        
+if len(valdict)>0:
+    print("Warning, not all values were checked. Unchecked: {}".format(" ".join(valdict.keys())))
+    ok=False
+
+if not ok:
+    sys.exit(1)

### 2024-05-13 scripts/spp_dcj.py
Fix more bugs/unknown vars.

@@ -388,7 +388,7 @@ def c36(G,tree_edge,out):
 
 def c37(G,tree_edge,out):
     for u,v,data in G.edges(data=True):
-        if data['type']==du.ETYPE_ADJ:
+        if data['type']!=du.ETYPE_ID:
             continue
         print("w{sep}{te}{sep}{u} - w{sep}{te}{sep}{v} = 0".format(
             sep=du.SEP,
@@ -443,10 +443,7 @@ def domains(graphs, out):
         for v,data in G.nodes(data=True):
             if data['type']==du.VTYPE_CAP:
                 continue
-            global_generals.add("g{sep}{v}".format(sep=du.SEP,v=v))
             print("0 <= y{sep}{te}{sep}{v} <= {v}".format(sep=du.SEP,te=te,v=v))
-    for gg in global_generals:
-        print("0 <= {gg} <= inf".format(gg=gg),file=out)
         
 
 
@@ -472,7 +469,7 @@ def variables(graphs, out):
             if data['type']==du.VTYPE_CAP:
                 continue
             print("w{sep}{te}{sep}{v}".format(sep=du.SEP,te=te,v=v),file=out)
-            global_generals.add("g{sep}{v}".format(sep=du.SEP,v=v))
+            
 
     for gg in global_generals:
         print(gg,file=out)
@@ -490,10 +487,12 @@ def variables(graphs, out):
         for v,data in G.nodes(data=True):
             print("z{sep}{te}{sep}{v}".format(sep=du.SEP,v=v,te=tree_edge),file=out)
             print("l{sep}{te}{sep}{v}".format(sep=du.SEP,v=v,te=tree_edge),file=out)
+            global_binaries.add("g{sep}{v}".format(sep=du.SEP,v=data['anc']))
             if data['type']!=du.VTYPE_CAP:
-                print("rab{sep}{te}{sep}{v}".format(sep=du.SEP,v=v,te=tree_edge),file=out)
+                if get_genome(G,v) == genomes[0]:
+                    print("rab{sep}{te}{sep}{v}".format(sep=du.SEP,v=v,te=tree_edge),file=out)
+                    print("rc{sep}{te}{sep}{v}".format(sep=du.SEP,te=tree_edge,v=v),file=out)
                 print("d{sep}{te}{sep}{v}".format(sep=du.SEP,te=tree_edge,v=v),file=out)
[…]

### 2024-05-13 docs/manuscript/main.tex
Write ILP description.

--- 2023-11-03 docs/manuscript/main.tex
@@
 The algorithm described in the following operates on two levels: 
-on the local level, a genome is derived from each GAG, constituting a set of linear or circular chromosomes. 
-On the global level, genomes are connected to each other along the branches of the phylogeny. Each branch gives rise to a pairwise comparison by means of the CFMRD. 
+on the global level, a genome is derived from each GAG, constituting a set of linear or circular chromosomes. 
+On the local level, genomes are connected to each other along the branches of the phylogeny. Each branch gives rise to a pairwise comparison by means of the CFMRD. 
 In doing so, the selection of adjacencies of a derived genome is propagated from across CFMRDs, thus ensuring global consistency.
 
+\paragraph{Global level.} 
+The global level deals with the setting of adjacencies or telomeres of (ancestral) genomes.
+For each extremity or pseudo-cap $\nu$, we determine its presence or absence with a binary variable $\ilpvar{g}_\nu$.
+We require each extremity to occur, except for pseudo-caps (see Constraints~\ref{c:cn}, \ref{c:cn_consistent}).
+Each extremity is required to be part of exactly one (possibly telomeric) adjacency (\ref{c:genome}), which ensures a properly linearized genome.
+
 \paragraph{Local level.}
-Constraint \ref{c:cn} implements the user-provided copy number requirement. 
-In doing so, we define the set of tail extremities of markers from family $i$ in GAG $\ggraph(\X)$ as $F_i = \{ v \mid v \in V: \family(v) = i \text{ and } \extf(v) = \et\}$ and iterate over the associated $\ilpvar{g}$ variables of its members to bound the number of active markers within the given range.
-The choice of counting tail extremities is arbitrary, however, the subsequent constraint (\ref{c:cn_consistent}) ensures that the assignment of $\ilpvar{g}$ variables for both extremities of a markers is consistent.
-At last, Constraint \ref{c:genome} makes sure that a valid genome is derived from the GAG when solving the ILP, by enforcing that each extremity node of $\ggraph(\X)$ is incident to at most one adjacency edge.
+The local level deals with each branch of the tree separately, making use of the CFMRD of the genome pair.
+Since this part is entirely local to the branch in question, we presume that each vertex $v_i$ of the CFMRD has a unique identifier among all other CFMRDs, making all its variables globally unique.
+In order to limit the range of the general variable $y_{v_i}$, we also assign each vertex a rank $i$ that is local and only unique within the specific CFMRD.
 
-\paragraph{Global level.}
+In order to compute decompositions of CFMRDs, we make use of a capping-free formulation for the computation of the pairwise DCJ indel distance~\cite{BOH-2024}.
+This formulation is based on the following distance formula,
 
-When comparing pairs of genomes within a CFMRD, we make use of a capping-free formulation for the computation of the pairwise DCJ indel distance~\cite{BOH-2023}.
+\begin{equation}
+	.... TODO.
+\end{equation}
+
+The formulation counts cycles $\ilpvar{c}_E$ as well as 6 different types of paths, namely $\ilpvar{p}^{AB},\ilpvar{p}^{Aa},\ilpvar{p}^{Ab},\ilpvar{p}^{ab},\ilpvar{p}^{Ba},\ilpvar{p}^{Bb}$, depending on whether a path ends in a pseudo-cap (uppercase letter) or a lava vertex (lowercase letter) and the genome in which it ends ($A,B$).
+Each counting variable $\ilpvar{p}^X$ is set by summing up binary report variables $\ilpvar{r}^X_v$ that are set to $1$ once per component on a specific vertex $v$ (see Constraints~\ref{c:beginsum} to~\ref{c:endsum} and~\ref{c:outsum}).
+These counters are then combined to the terms of the formula in Constraints~\ref{c:begingeq} to~\ref{c:endgeq} and \ref{c:beginform} to~\ref{c:endform}.
+The constraints for ensuring the reporting variables being set correctly can be found in Tables~\ref{tab:slmcons}, \ref{tab:regv} and~\ref{tab:pcaps}.
+For a complete description of the ILP the interested reader is referred to~\cite{BOH-2024}.
+
+We make only few major changes in our local section w.r.t.\ the ILP described in~\cite{BOH-2024}.
+Firstly, we determine whether an adjacency edge $e$ is set ($x_e=1$) by ``inheriting'' this value from the global setting (see~\ref{c:inheritadj}) of the corresponding adjacency.
+Secondly, we only allow vertices that are part of the genome ($\ilpvar{g}_v=1$) to contribute to the count of components that decrease the formula ($\ilpvar{z}_v=1$), see~\ref{c:inheritz}.
+Finally, due to the fact that ancestral genomes are degenerate, the number of possible circular singletons is much higher and listing all candidates might lead to a combinatorial explosion on some input data.
+We therefore use a new technique for counting circular singletons without listing all candidates, the constraints of which are listed in Table~\ref{tab:csreport}.
+
+A circular singleton manifests in the graph as a cycle of alternating adjacency and indel edges.
+The idea of the technique is to have a general integer variable $\ilpvar{w}$ that is required to increase at each adjacency edge in a walk of the cycle.
+There must then be one point in the walk in which it decreases again.
+Detecting this, one can then report a circular singleton.
+For this to work, the walk needs a direction.
+This is accomplished by annotating the vertices with a binary variable $\ilpvar{d}_v$ that ``flips'' across each pair of connected vertices (see~\ref{c:dflip}).
+We then require $\ilpvar{w}$ to be the same for vertices connected by an indel edge (see~\ref{c:weq}) and for it to increase by $1$ in the direction of the vertex that has $\ilpvar{d}_v=1$ (see~\ref{c:winc}).
+We require this except when vertices are not connected $(1-\ilpvar{x}_{uv})=0$ or when reporting a circular singleton ($\ilpvar{r}_v=1$ or $\ilpvar{r}_u=1$).
+In this case, the constraint is automatically fulfilled by adding the maximum length of circular singletons $K$ to the left hand side of the inequation.
 

### 2024-05-13 scripts/data_utils.py
fixed warning on regexp string

@@ -51,8 +51,8 @@ TRUE_ADJ_WEIGHT  = 1
 
 DEFAULT_GENE_FAM_SEP = '_'
 
-PAT_ADJ = re.compile('^(\w+)@([0-9_]+)$')
-PAT_MATCHED_EDGE = re.compile('^x(\d+)_(\d+)([^0-9 \t]*) 1\s*$')
+PAT_ADJ = re.compile(r'^(\w+)@([0-9_]+)$')
+PAT_MATCHED_EDGE = re.compile(r'^x(\d+)_(\d+)([^0-9 \t]*) 1\s*$')
 
 
 def complement_id(v):

### 2024-05-15 scripts/data_utils.py
Implement CS filter.

@@ -712,8 +712,8 @@ def _find_end_pairs(G, gName1, gName2):
 
 
 def checkGraph(G,cf=False):
-    for v,data in G.nodes(data=True):
-        print(v,",".join(["{}={}".format(k,x) for k,x in data.items()]),file=sys.stderr)
+    #for v,data in G.nodes(data=True):
+    #    print(v,",".join(["{}={}".format(k,x) for k,x in data.items()]),file=sys.stderr)
     for u, v, in G.edges():
         if u == v:
             raise Exception(f'node {v} is connected to itself')
@@ -783,6 +783,24 @@ def identifyCircularSingletonCandidates(G):
     return res
 
 
+def annotate_v_circ_sing(G):
+    #Find all vertices that could be in a circular singleton
+    F = G.copy()
+    rmedges= [(u,v,k) for u,v,k,etype in G.edges(keys=True,data='type') if etype==ETYPE_EXTR]
+    #print("rmedge: {}".format(rmedges),file=sys.stderr)
+    F.remove_edges_from(rmedges)
+
+    for c in nx.connected_components(F):
+        S=F.subgraph(c)
+        try:
+            nx.find_cycle(S)
+            #print("CS candidate: {}".format(c),file=sys.stderr)
+            for v in c:
+                G.nodes[v]['cscandidate']=True
+        except nx.NetworkXNoCycle:
+            pass
+
+
 def rotateToMin(path):
     m = min((path[i] for i in range(0, len(path), 2)))
     i = path.index(m)
@@ -829,6 +847,7 @@ def _constructRDExtremityEdges(G, gName1, gName2, genes, fam2genes1,
         for i, gName in enumerate((gName1, gName2)):
             if len(fam2genes[i].get(fam, ())) > len(fam2genes[i-1].get(fam, \
                     ())):
+                #print("Fam {} in genome {} overrepresented. Adding indel edges...".format(fam,gName),file=sys.stderr)
                 for gene in fam2genes[i][fam]:
                     idh = extremityIdManager.getId((gName, (gene, EXTR_HEAD)))
                     idt = extremityIdManager.getId((gName, (gene, EXTR_TAIL)))
@@ -902,6 +921,7 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
         
         res['graphs'][(child, parent)] = G
         res['siblings'][(child, parent)] = siblings
+        annotate_v_circ_sing(G)
 
 
     # create caps at last, assigning them the highest IDs

### 2024-05-15 scripts/spp_dcj.py

@@ -367,7 +367,7 @@ def pcap_reporting(G,tree_edge,genomes,out):
 
 def c36(G,tree_edge,out):
     for u,v,data in G.edges(data=True):
-        if du.VTYPE_CAP in [G.nodes[x]['type'] for x in [u,v]]:
+        if du.VTYPE_CAP in [G.nodes[x]['type'] for x in [u,v]] or not G.nodes[u].get('cscandidate',False):
             continue
         if data['type']==du.ETYPE_EXTR:
             continue
@@ -388,7 +388,7 @@ def c36(G,tree_edge,out):
 
 def c37(G,tree_edge,out):
     for u,v,data in G.edges(data=True):
-        if data['type']!=du.ETYPE_ID:
+        if data['type']!=du.ETYPE_ID or not G.nodes[u].get('cscandidate',False):
             continue
         print("w{sep}{te}{sep}{u} - w{sep}{te}{sep}{v} = 0".format(
             sep=du.SEP,
@@ -399,7 +399,7 @@ def c37(G,tree_edge,out):
 
 def c38(G,tree_edge,out,max_circ_len):
     for u_,v_,data in G.edges(data=True):
-        if data['type']!=du.ETYPE_ADJ or du.VTYPE_CAP in [G.nodes[x]['type'] for x in [u_,v_]]:
+        if data['type']!=du.ETYPE_ADJ or du.VTYPE_CAP in [G.nodes[x]['type'] for x in [u_,v_]] or not G.nodes[u_].get('cscandidate',False):
             continue
         for u,v in [(u_,v_),(v_,u_)]:
             print("w{sep}{te}{sep}{u} - w{sep}{te}{sep}{v} + d{sep}{te}{sep}{v} - d{sep}{te}{sep}{u} + {K} x{sep}{te}{sep}{e} - {K} rs{sep}{te}{sep}{u} - {K} rs{sep}{te}{sep}{v} <= {K}".format(
@@ -466,7 +466,7 @@ def variables(graphs, out):
         print("q{sep}{e}".format(sep=du.SEP,e=te))
         for v,data in G.nodes(data=True):
             print("y{sep}{te}{sep}{v}".format(sep=du.SEP,te=te,v=v),file=out)
-            if data['type']==du.VTYPE_CAP:
+            if data['type']==du.VTYPE_CAP or not G.nodes[v].get('cscandidate',False):
                 continue
             print("w{sep}{te}{sep}{v}".format(sep=du.SEP,te=te,v=v),file=out)
             
[…]

### 2024-05-15 scripts/data_utils.py
Implement selecting old circular singleton constraint when few singletons exist.

@@ -747,7 +747,10 @@ def checkGraph(G,cf=False):
                     'to an extremity or indel edge')
 
 
-def identifyCircularSingletonCandidates(G):
+class TooManyCSException(Exception):
+    pass
+
+def identifyCircularSingletonCandidates(G,max_number=None):
     """ finds all components that can be circular singletons """
 
     res = dict()
@@ -777,9 +780,14 @@ def identifyCircularSingletonCandidates(G):
                             ppath = rotateToMin(path + (data, ))
                             vpath = tuple((ppath[i] for i in range(0,
                                 len(ppath), 2)))
+                            #TODO: Tell dany I found a bug here!
+                            vpath=canonicizePath(vpath)
                             epath = tuple((ppath[i] for i in range(1,
                                 len(ppath), 2)))
                             res[vpath] = epath
+                            if max_number is not None:
+                                if len(res)>max_number:
+                                    raise TooManyCSException("Number of CS is over maximum {}.".format(max_number))
     return res
 
 
@@ -805,6 +813,19 @@ def rotateToMin(path):
     m = min((path[i] for i in range(0, len(path), 2)))
     i = path.index(m)
     return path[i:] + path[:i]
+
+def canonicizePath(path):
+    m = min(path)
+    i = path.index(m)
+    path = path[i:] + path[:i]
+    if len(path) > 2:
+        if path[1] <= path[-1]:
+            return path
+        else:
+            return (path[0],) + path[-1:0:-1]
+    return path
+        
+
 
 
 def _constructRDExtremityEdges(G, gName1, gName2, genes, fam2genes1,

### 2024-05-15 scripts/spp_dcj.py

@@ -40,11 +40,12 @@ def objective(graphs, alpha, out):
         
 
 
+
 #
 # ILP CONSTRAINTS
 #
 
-def constraints(graphs, siblings, out):
+def constraints(graphs, siblings,circ_singletons, out):
     out.write('subject to\n')
     for i, ((child, parent), G) in enumerate(sorted(graphs.items())):
 
@@ -56,7 +57,12 @@ def constraints(graphs, siblings, out):
         slm_constraints(G,i,siblings,out)
         regular_reporting(G,i,genomes,out)
         pcap_reporting(G,i,genomes,out)
-        cs_constraints(G,i,out,max_circ_len=len(G.nodes())/2)
+        if (child,parent) in circ_singletons:
+            c19(G,i,out,candidates=circ_singletons[(child,parent)])
+            manual_cs_constraints(circ_singletons[(child,parent)],i,out)
+        else:
+            c19(G,i,out)
+            cs_constraints(G,i,out,max_circ_len=len(G.nodes())/2)
 
     out.write('\n')
 
@@ -129,8 +135,14 @@ def c14toc17(G,tree_edge,out):
             print("pAB{ind}{sep}{te} - p{tel}{ind}{sep}{te} >= 0".format(ind=indel,tel=tel,te=tree_edge,sep=du.SEP),file=out)
 c18 = lambda G,tree_edge,genomes,out: summary_constraint('AB',lambda G,v: get_genome(G,v)==genomes[0] and G.nodes[v]['type']==du.VTYPE_CAP,G,tree_edge,out)
 
-def c19(G,tree_edge,out):
-    sm = ['rs{sep}{te}{sep}{v}'.format(te=tree_edge,sep=du.SEP,v=v) for v,data in G.nodes(data=True) if data['type'] != du.VTYPE_CAP]
+def c19(G,tree_edge,out,candidates=None):
+    if candidates is None:
+        sm = ['rs{sep}{te}{sep}{v}'.format(te=tree_edge,sep=du.SEP,v=v) for v,data in G.nodes(data=True) if data['type'] != du.VTYPE_CAP and data.get('cscandidate',False)]
+        
+    else:
+        LOG.info("manual candidates: {}".format(candidates))
+        sm = ['rms{sep}{te}{sep}{j}'.format(sep=du.SEP,te=tree_edge,j=j) for j in range(len(candidates))]
+        LOG.info("sm: {}".format(sm))
[…]

### 2024-05-15 scripts/spp_dcj.py
Add option for circular singleton handling

@@ -622,6 +622,7 @@ if __name__ == '__main__':
     parser.add_argument('-b', '--beta', default = -1, type=float,
             help='linear weighting factor for telomeric adjacencies;' + \
                     'if beta < 0, then beta is set to 1/2 * alpha')
+    parser.add_argument('--set-circ-sing-handling',choices=['adaptive','enumerate','barbershop'],default='adaptive')
     parser.add_argument('-s', '--separator', default = du.DEFAULT_GENE_FAM_SEP, \
             help='Separator of in gene names to split <family ID> and ' +
                     '<uniquifying identifier> in adjacencies file')
@@ -709,8 +710,16 @@ if __name__ == '__main__':
 
     circ_singletons = dict()
     for ident, G in graphs.items():
-        max_tolerable = len([u for u in G.nodes() if G.nodes[u].get('cscandidate',False)])
-        if max_tolerable > 0:
+        if args.set_circ_sing_handling=='adaptive':
+            max_tolerable = len([u for u in G.nodes() if G.nodes[u].get('cscandidate',False)])
+        elif args.set_circ_sing_handling=='enumerate':
+            max_tolerable = None
+        else:
+            max_tolerable=0
+        if max_tolerable==None:
+            circ_singletons[ident] = du.identifyCircularSingletonCandidates(G)
+            LOG.info(f'identified {len(circ_singletons[ident])} circular singleton candidates')
+        elif max_tolerable > 0:
             try:
                 circ_singletons[ident] = du.identifyCircularSingletonCandidates(G,max_number=max_tolerable)
                 LOG.info(f'identified {len(circ_singletons[ident])} circular singleton candidates')

### 2024-05-15 scripts/spp_dcj.py
Remove beta, add default telomere weight.

@@ -533,7 +533,7 @@ def variables(graphs,circ_sings, out):
     print('\n')
 
 
-def identifyCandidateTelomeres(candidateAdjacencies, weightThreshold, dont_add=False):
+def identifyCandidateTelomeres(candidateAdjacencies, telomere_default_weight, dont_add=False):
 
     res = dict()
     weights = candidateAdjacencies['weights']
@@ -579,6 +579,7 @@ def identifyCandidateTelomeres(candidateAdjacencies, weightThreshold, dont_add=F
                         t = f't_{g}_{extr}'
                         telomeres.add(t)
                         adjs.append(((g, extr), (t, 'o')))
+                        candidateAdjacencies['weights'][((g, extr), (t, 'o'))]=telomere_default_weight
 
 #        genes_edg = [((g, du.EXTR_HEAD), (g, du.EXTR_TAIL)) for g in genes]
 #        if species == 'n3':
@@ -619,9 +620,10 @@ if __name__ == '__main__':
             help='linear weighting factor for adjacency weights vs DCJ ' + \
                     'indel distance (alpha = 1 => maximize only DCJ indel ' + \
                     'distance)')
-    parser.add_argument('-b', '--beta', default = -1, type=float,
-            help='linear weighting factor for telomeric adjacencies;' + \
-                    'if beta < 0, then beta is set to 1/2 * alpha')
+    #parser.add_argument('-b', '--beta', default = -1, type=float,
+    #        help='Default weight for added telomeres. Has no effect if -t is used.')
+    parser.add_argument('-def-telomere-weight', '--dtw', default = -1, type=float,
+            help='Default weight for added telomeres. Has no effect if -t is used. For most use cases this should be <= 0.')
     parser.add_argument('--set-circ-sing-handling',choices=['adaptive','enumerate','barbershop'],default='adaptive')
     parser.add_argument('-s', '--separator', default = du.DEFAULT_GENE_FAM_SEP, \
             help='Separator of in gene names to split <family ID> and ' +
@@ -634,9 +636,6 @@ if __name__ == '__main__':
     ch.setFormatter(logging.Formatter('%(levelname)s\t%(asctime)s\t%(message)s'))
     LOG.addHandler(ch)
 
-    beta = args.beta
-    if beta < 0:
-        beta = args.alpha * 0.5
 
     # load & process input data
     LOG.info('loading species tree from {}'.format(args.tree.name))
[…]

### 2024-05-15 scripts/data_utils.py
Implement all telomere feature

@@ -711,7 +711,7 @@ def _find_end_pairs(G, gName1, gName2):
     return {(u, v, Arun, Brun) for (u, v), (Arun, Brun) in res.items()}
 
 
-def checkGraph(G,cf=False):
+def checkGraph(G,cf=False,checkForAllTels=False):
     #for v,data in G.nodes(data=True):
     #    print(v,",".join(["{}={}".format(k,x) for k,x in data.items()]),file=sys.stderr)
     for u, v, in G.edges():
@@ -734,7 +734,12 @@ def checkGraph(G,cf=False):
         if vdata['id'][1][1] not in {EXTR_HEAD, EXTR_TAIL, EXTR_CAP}:
             raise Exception(f'node {v} {G.nodes[v]["id"]} has malformed ' + \
                     'extremity')
-
+        # if vdata['id'][1][1] != EXTR_CAP and checkForAllTels:
+        #     has_cap=False
+        #     for u in G.neighbors(v):
+        #         has_cap=has_cap or G.nodes[u]['type']==EXTR_CAP
+        #     if not has_cap:
+        #         raise Exception("Looked for caps, but node {} ({}) is lacking one.".format(v,vdata))
         for u in G.neighbors(v):
             for data in G[u][v].values():
                 hasAdj |= data['type'] == ETYPE_ADJ
@@ -926,7 +931,7 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
 
     for child, parent in tree:
         G = nx.MultiGraph()
-        max_tels = 2*sum([len(genes[gnm]) for gnm in [child,parent]])
+        max_tels = len(candidateTelomeres[child]) + len(candidateTelomeres[parent]) +2 #2*sum([len(genes[gnm]) for gnm in [child,parent]])
         localIdManager = IdManager(max_tels)
         for gName in (child, parent):
             _constructRDNodes(G, gName, genes[gName], extremityIdManager,localIdManager)

### 2024-05-15 scripts/spp_dcj.py

@@ -533,7 +533,7 @@ def variables(graphs,circ_sings, out):
     print('\n')
 
 
-def identifyCandidateTelomeres(candidateAdjacencies, telomere_default_weight, dont_add=False):
+def identifyCandidateTelomeres(candidateAdjacencies, telomere_default_weight,leaves, dont_add=False,addToAll=False):
 
     res = dict()
     weights = candidateAdjacencies['weights']
@@ -574,8 +574,11 @@ def identifyCandidateTelomeres(candidateAdjacencies, telomere_default_weight, do
                 # - linear/circular or
                 # - fully connected
                 # - even (in terms of #nodes)
-                if degs.difference((1, 2)) and degs.difference((len(C)-1,)) or len(C) % 2:
+                if degs.difference((1, 2)) and degs.difference((len(C)-1,)) or len(C) % 2 or (addToAll and species not in leaves):
                     for g, extr in C:
+                        #skip already existing telomeres
+                        if extr=='o':
+                            continue
                         t = f't_{g}_{extr}'
                         telomeres.add(t)
                         adjs.append(((g, extr), (t, 'o')))
@@ -613,7 +616,9 @@ if __name__ == '__main__':
     parser.add_argument('candidateAdjacencies', type=open,
             help='candidate adjacencies of the genomes in the phylogeny')
     parser.add_argument('-t', '--no_telomeres', action='store_true',
-            help='don\'t add any additional telomeres')
+            help='don\'t add any additional telomeres. Overrides -at.')
+    parser.add_argument('-l', '--all_telomeres', action='store_true',
+            help='Add additional telomeres to all nodes.')
     parser.add_argument('-m', '--output_id_mapping', type=FileType('w'),
             help='writs a table with ID-to-gene extremity mapping')
     parser.add_argument('-a', '--alpha', default = 0.5, type=float,
@@ -622,7 +627,7 @@ if __name__ == '__main__':
                     'distance)')
     #parser.add_argument('-b', '--beta', default = -1, type=float,
     #        help='Default weight for added telomeres. Has no effect if -t is used.')
-    parser.add_argument('-def-telomere-weight', '--dtw', default = -1, type=float,
+    parser.add_argument('--def-telomere-weight', '-w', default = -1, type=float,
[…]

### 2024-05-15 scripts/data_utils.py
Local changes by dany

@@ -129,8 +129,7 @@ def parseAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
     delimiter = '\t'
     resAdjacencies = defaultdict(list)
     resGenes       = defaultdict(list)
-    resWeights     = {}
-    resPenalities  = {}
+    resWeights     = defaultdict(dict)
     for line in csv.reader(data, delimiter = delimiter):
         if line[0][0] != headerMark:
             species  = line[0]
@@ -144,11 +143,7 @@ def parseAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
                         f'feasible, skipping {line[1]}:{line[2]}-{line[4]}:{line[5]}',
                         file=stderr)
             else:
-                addAdjacency(ext1,ext2,weight,resAdjacencies[species],resWeights,resGenes[species])
-
-            # optional penality
-            if len(line) > 7 and line[7]:
-                resPenalities[ext1 > ext2 and (ext2, ext1) or (ext1, ext2)] = float(line[7])
+                addAdjacency(ext1,ext2,weight,resAdjacencies[species],resWeights[species],resGenes[species])
 
     speciesList = list(resAdjacencies.keys())
     for species in speciesList:
@@ -157,7 +152,7 @@ def parseAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
 
     return {'species':speciesList, 'genes':resGenes,
             'adjacencies':resAdjacencies, 'weights':resWeights,
-            'families': resFamilies, 'penalities': resPenalities}
+            'families': resFamilies}
 
 
 def parseCandidateAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
@@ -172,7 +167,7 @@ def parseCandidateAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
     delimiter = ' '
     resAdjacencies = defaultdict(list)
     resGenes       = defaultdict(list)
-    resWeights     = {}
+    resWeights     = defaultdict(dict)
     for line in csv.reader(data, delimiter = delimiter):
         if line[0][0] != headerMark:
             species = line[0]
@@ -192,7 +187,7 @@ def parseCandidateAdjacencies(data, sep=DEFAULT_GENE_FAM_SEP):
             weight  = float(line[5])
             ext1 = (gene1,SIGN2EXT_1[sign1])
             ext2 = (gene2,SIGN2EXT_2[sign2])
[…]

### 2024-05-15 scripts/spp_dcj.py

@@ -631,13 +631,12 @@ if __name__ == '__main__':
     genes = candidateAdjacencies['genes']
     adjacencies = candidateAdjacencies['adjacencies']
     weights = candidateAdjacencies['weights']
-    penalities = candidateAdjacencies['penalities']
 
     global_ext2id = du.IdManager(0,is_cap=lambda x: False)
     LOG.info(('constructing relational diagrams for all {} branches of ' + \
             'the tree').format(len(speciesTree)))
     relationalDiagrams = du.constructRelationalDiagrams(speciesTree,
-            adjacencies, telomeres, weights, penalities, genes, global_ext2id,
+            adjacencies, telomeres, weights, genes, global_ext2id,
             sep=args.separator)
 
     graphs = relationalDiagrams['graphs']

### 2024-05-15 tools/recPhyloXML/ReconciledTreeIO.py

@@ -144,7 +144,7 @@ class recPhyloXML_parser:
         if not self.isOfTag(element, TAG):
             raise Exception('BadTagException. The element is of tag ' + element.tag + " instead of " + TAG + "." )
 
-        children = element.getchildren()
+        children = list(element)
 
         RTL = ReconciledTreeList()
 
@@ -179,7 +179,7 @@ class recPhyloXML_parser:
         if not self.isOfTag(element, TAG):
             raise Exception('BadTagException. The element is of tag ' + element.tag + " instead of " + TAG + "." )
 
-        children = element.getchildren()
+        children = list(element)
 
         node = None
 
@@ -211,7 +211,7 @@ class recPhyloXML_parser:
         if not self.isOfTag(element, TAG):
             raise Exception('BadTagException. The element is of tag ' + element.tag + " instead of " + TAG + "." )
 
-        children = element.getchildren()
+        children = list(element)
 
         node = None
 
@@ -246,7 +246,7 @@ class recPhyloXML_parser:
         if not self.isOfTag(element, TAG):
             raise Exception('BadTagException. The element is of tag ' + element.tag + " instead of " + TAG + "." )
 
-        children = element.getchildren()
+        children = list(element)
 
         node = None
 
@@ -299,7 +299,7 @@ class recPhyloXML_parser:
         if not self.isOfTag(element, TAG):
             raise Exception('BadTagException. The element is of tag ' + element.tag + " instead of " + TAG + "." )
 
-        children = element.getchildren()
+        children = list(element)
 
         name = None
         childrenNodes = []
@@ -369,7 +369,7 @@ class recPhyloXML_parser:
         if not self.isOfTag(element, TAG):
             raise Exception('BadTagException. The element is of tag ' + element.tag + " instead of " + TAG + "." )
 
-        children = element.getchildren()
+        children = list(element)
 
         events = []
 

### 2025-01-15 scripts/spp_dcj.py
Untested Warm start and fix tests

@@ -465,6 +465,244 @@ def domains(graphs, out):
         
 
 
+def create_sibmap(G,siblings):
+    id_to_edges = []
+    sibmap=dict()
+    for u,v,data in G.edges(data=True):
+        if data['etype']==du.ETYPE_EXTR:
+            id_to_edges[data['id']]=(u,v)
+    for siba,sibb in siblings:
+        if not siba in id_to_edges or not sibb in id_to_edges:
+            continue
+        x,y=id_to_edges[siba]
+        w,z =id_to_edges[sibb]
+        for u,v in [(x,y),(y,x)]:
+            sibmap[(u,v)]=(w,z)
+        for u,v in [(w,z),(z,w)]:
+            sibmap[(u,v)]=(x,y)
+    return sibmap
+
+
+            
+
+def warm_start_decomposition(graphs,siblings):
+    create_max_weight(graphs)
+    for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
+        sibmap=create_sibmap(G,siblings)
+        heuristic_partial_matching(G,sibmap)
+        fill_matching_greedy(G,sibmap)
+
+
+def create_max_weight(graphs):
+    root_candidates = dict()
+    matchings = dict()
+    for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
+        root_candidates[parent]=child
+        root_candidates.pop(child,None)
+        matchings[child]=get_max_match(child, G)
+    assert(len(root_candidates)==1)
+    parent,child = root_candidates.items()[0]
+    G=graphs[child]
+    matchings[parent]=get_max_match(parent,G)
+    for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
+        set_based_on_matching(G,matchings[child],child)
+        set_based_on_matching(G,matchings[parent],parent)
+    
+
+def get_anc_map(g):
+    return dict(((anc,v) for v,anc in g.nodes(data='anc')))
+
+
+def set_based_on_matching(G,mtch,genome):
+    local = extract_local_graph(genome, G)
+    ancmap = get_anc_map(local)
+    for u_,v_ in mtch:
+        u,v = ancmap[u_],ancmap[v_]
+        G[u][v]['is_set']=True
+    for u,v in local.edges():
+        if not 'is_set' in G[u][v]:
+            G[u][v]=False
+
+def get_max_match(genome, G):
+    local = extract_local_graph(genome, G)
[…]

### 2025-01-17 scripts/spp_dcj.py
Preliminary version of warm start, still untested

@@ -466,11 +466,12 @@ def domains(graphs, out):
 
 
 def create_sibmap(G,siblings):
-    id_to_edges = []
+    id_to_edges = dict()
     sibmap=dict()
     for u,v,data in G.edges(data=True):
-        if data['etype']==du.ETYPE_EXTR:
+        if data['type']==du.ETYPE_EXTR:
             id_to_edges[data['id']]=(u,v)
+    #print(siblings,file=stderr)
     for siba,sibb in siblings:
         if not siba in id_to_edges or not sibb in id_to_edges:
             continue
@@ -488,7 +489,7 @@ def create_sibmap(G,siblings):
 def warm_start_decomposition(graphs,siblings):
     create_max_weight(graphs)
     for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
-        sibmap=create_sibmap(G,siblings)
+        sibmap=create_sibmap(G,siblings[(child,parent)])
         heuristic_partial_matching(G,sibmap)
         fill_matching_greedy(G,sibmap)
 
@@ -496,17 +497,19 @@ def warm_start_decomposition(graphs,siblings):
 def create_max_weight(graphs):
     root_candidates = dict()
     matchings = dict()
+    
     for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
         root_candidates[parent]=child
         root_candidates.pop(child,None)
         matchings[child]=get_max_match(child, G)
     assert(len(root_candidates)==1)
-    parent,child = root_candidates.items()[0]
-    G=graphs[child]
+    parent,child = list(root_candidates.items())[0]
+    G=graphs[(child,parent)]
     matchings[parent]=get_max_match(parent,G)
     for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
         set_based_on_matching(G,matchings[child],child)
         set_based_on_matching(G,matchings[parent],parent)
+        #print(G.edges(data=True),file=stderr)
     
 
 def get_anc_map(g):
@@ -518,17 +521,49 @@ def set_based_on_matching(G,mtch,genome):
     ancmap = get_anc_map(local)
     for u_,v_ in mtch:
         u,v = ancmap[u_],ancmap[v_]
-        G[u][v]['is_set']=True
-    for u,v in local.edges():
-        if not 'is_set' in G[u][v]:
-            G[u][v]=False
+        candidates = [k for k in G[u][v] if G[u][v][k]['type']==du.ETYPE_ADJ]
[…]

### 2025-01-29 scripts/spp_dcj.py
Fix warm start

@@ -101,7 +101,6 @@ def constraints(graphs,families,fam_bounds, siblings,circ_singletons, out,lower_
             cs_constraints(G,i,out,max_circ_len=len(G.nodes())/2)
     LOG.info("Writing lower bound constraints.")
     cp_tree=child_parent_tree(graphs)
-    print(cp_tree,file=stderr)
     for a in lower_bound_mat:
         for b in lower_bound_mat[a]:
             lb = lower_bound_mat[a][b]
@@ -143,7 +142,6 @@ def c01(G,families,fam_bounds,out):
     assert(len(id_to_v)==0)
     for u,v in ht_pairs:
         f=du.nodeGetFamily(G,u)
-        print(f,file=stderr)
         assert(f==du.nodeGetFamily(G,v))
         gnm = get_genome(G,v)
         if len(families[gnm][f])==fam_bounds[gnm][f][0]:
@@ -609,7 +607,6 @@ def create_sibmap(G,siblings):
 
 def warm_start_decomposition(graphs,fam_bounds,families,siblings):
     not_full_fams = du.getNotFullFamilies(fam_bounds,families)
-    print(not_full_fams,file=stderr)
     create_max_weight(graphs,not_full_fams,siblings)
     for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
         sibmap=create_sibmap(G,siblings[(child,parent)])
@@ -673,15 +670,11 @@ def create_max_weight(graphs,not_full_fams,siblings):
     superfluous_anc = dict()
     for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
         sibmap=create_sibmap(G,siblings[(child,parent)])
-        print(parent,child, file=stderr)
         root_candidates[parent]=child
-        print(root_candidates,file=stderr)
         not_root.add(child)
-        print(root_candidates,file=stderr)
         superfluous_anc[child] = get_superfluous_ancs(G,sibmap,not_full_fams,child)
         annotate_superfluous(G,superfluous_anc[child])
         matchings[child]=get_max_match(child, G)
-    print(root_candidates,file=stderr)
     for x in not_root:
         root_candidates.pop(x,None)
     assert(len(root_candidates)==1)
@@ -1027,10 +1020,8 @@ def heuristic_partial_matching(G,sibmap,max_iter=None):
         etype = invert_etype(etype)
         rm_adj = set()
         changed = False
-        print('-----',file=stderr)
[…]

### 2025-01-30 scripts/spp_dcj.py
init solution for circ sings - partially implemented

@@ -1303,10 +1303,28 @@ def sol_from_decomposition(graphs,circ_singletons,alpha,out):
                 counts['s']+=is_set
         else:
             #TODO: implement other circ sing counting
-            for v,data in G.nodes(data=True):
+            for x, data in G.nodes(data=True):
                 if data.get('cscandidate',False):
                     counts['s']+=1
-            pass
+            #local=G.copy()
+            #for u,v,k,data in G.edges(keys=True,data=True):
+            #    if data['type']==du.ETYPE_EXTR or not data.get('is_set',False):
+            #        local.remove_edge(u,v,k)
+            #for comp in nx.connected_components(local):
+            #    p_ends = []
+            #    for v in comp:
+            #        dg = local.degree[v]
+            #        assert(dg <=2)
+            #        if dg==1:
+            #            p_ends.append(v)
+            #    assert(len(p_ends)<=2)
+            #    if len(p_ends)==1:
+            #        continue
+            #    if len(p_ends)==0:
+            #        for x in nx.edge_dfs(list(comp)[0]):
+            #            pass
+                    
+
         for u,data in G.nodes(data=True):
             ancvar = "g{sep}{anc}".format(sep=du.SEP,anc=data['anc'])
             acands = [v for v in G[u] for k in G[u][v] if G[u][v][k]['type']==du.ETYPE_ADJ and G[u][v][k].get('is_set',False)]

### 2025-01-30 scripts/compute_initial.py
Refactoring to have initial solution script independent of ding

@@ -1,0 +1,91 @@
+from argparse import ArgumentParser
+import data_utils as du
+
+def main():
+    parser = ArgumentParser()
+    parser.add_argument('tree', type=open,
+            help='phylogenetic tree as parent-child relation table')
+    parser.add_argument('candidateAdjacencies', type=open,
+            help='candidate adjacencies of the genomes in the phylogeny')
+    parser.add_argument('idmap',help="Id map file (output of the main script)",type=open)
+    parser.add_argument('warm_start_sol',help="File to write the heuristic solution to.")
+    parser.add_argument('-t', '--no_telomeres', action='store_true',
+            help='don\'t add any additional telomeres. Overrides -at.')
+    parser.add_argument('-l', '--all_telomeres', action='store_true',
+            help='Add additional telomeres to all nodes.')
+    parser.add_argument('--family-bounds','-fmb',type=open)
+    parser.add_argument('--set-circ-sing-handling',choices=['adaptive','enumerate','barbershop'],default='adaptive')
+    parser.add_argument('-s', '--separator', default = du.DEFAULT_GENE_FAM_SEP, \
+            help='Separator of in gene names to split <family ID> and ' +
+                    '<uniquifying identifier> in adjacencies file')
+    parser.add_argument('-a', '--alpha', default = 0.5, type=float,
+            help='linear weighting factor for adjacency weights vs DCJ ' + \
+                    'indel distance (alpha = 1 => maximize only DCJ indel ' + \
+                    'distance)')
+    parser.add_argument('--def-telomere-weight', '-w', default = 0, type=float,
+            help='Default weight for added telomeres. Has no effect if -t is used. For most use cases this should be <= 0.')
+    parser.add_argument('-b', '--beta', type=float,
+            help='Backwards compatible beta parameter from v1. Telomeres will be re-weighted and the ILP scaled to be equivalent to v1.')
+
+    args = parser.parse_args()
+    glob_table, loc_tables = du.parse_id_file(args.idmap)
+    global_ext2id = du.IdManager(0,is_cap=lambda x: False,initial_table=glob_table)
+    speciesTree = du.parseTree(args.tree)
+
+
+    candidateAdjacencies = du.parseAdjacencies(args.candidateAdjacencies,
+                                               sep=args.separator)
[…]

### 2025-01-30 scripts/data_utils.py

@@ -3,7 +3,9 @@
 # import from built-in packages
 from argparse import ArgumentParser, ArgumentDefaultsHelpFormatter as ADHF, FileType
 from collections import defaultdict, deque
+from functools import reduce
 from itertools import product, chain, combinations
+from math import ceil
 from sys import stderr
 import random
 import csv
@@ -14,6 +16,8 @@ import sys
 
 # import from third-party packages
 import networkx as nx
+
+
 
 #
 # global variables
@@ -932,15 +936,18 @@ def getIncidentAdjacencyEdges(G, v):
 
 def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
         candidateWeights, genes, extremityIdManager,fam_bounds=dict(),
-        sep=DEFAULT_GENE_FAM_SEP):
+        sep=DEFAULT_GENE_FAM_SEP,loc_manager_tables=dict()):
     ''' constructs for each edge of the tree a relational diagram of the
     adjacent genomes'''
 
-    res = dict((('graphs', dict()), ('siblings', dict())))
+    res = dict((('graphs', dict()), ('siblings', dict()),('idmanagers',dict())))
     for child, parent in tree:
         G = nx.MultiGraph()
         max_tels = len(candidateTelomeres[child]) + len(candidateTelomeres[parent]) +2 #2*sum([len(genes[gnm]) for gnm in [child,parent]])
-        localIdManager = IdManager(max_tels)
+        if (child,parent) in loc_manager_tables:
+            localIdManager=IdManager(max_tels,initial_table=loc_manager_tables[(child,parent)])
+        else:
+            localIdManager = IdManager(max_tels)
         for gName in (child, parent):
             _constructRDNodes(G, gName, genes[gName], extremityIdManager,localIdManager)
                 
@@ -956,6 +963,7 @@ def constructRelationalDiagrams(tree, candidateAdjacencies, candidateTelomeres,
         
         res['graphs'][(child, parent)] = G
         res['siblings'][(child, parent)] = siblings
+        res['idmanagers'][(child,parent)]=localIdManager
         annotate_v_circ_sing(G)
 
 
@@ -1009,14 +1017,17 @@ def getNotFullFamilies(fam_bounds,families):
 
 class IdManager(object):
 
-    def __init__(self,cap_limit,is_cap = lambda x : x[1][1]=='o'):
[…]

### 2025-01-30 scripts/spp_dcj.py

@@ -5,15 +5,12 @@ from argparse import ArgumentParser, ArgumentDefaultsHelpFormatter as ADHF, \
         FileType
 from sys import stdout, stderr, exit
 from itertools import product, combinations, chain, repeat
-from functools import reduce
 from collections import defaultdict
-from math import ceil
 import logging
 import csv
 from math import exp
 
 # import from third-party packages
-import networkx as nx
 
 # import from own packages
 import data_utils as du
@@ -583,905 +580,8 @@ def domains(graphs, out):
         
 
 
-def create_sibmap(G,siblings):
-    id_to_edges = dict()
-    sibmap=dict()
-    for u,v,data in G.edges(data=True):
-        if data['type']==du.ETYPE_EXTR:
-            id_to_edges[data['id']]=(u,v)
-    #print(siblings,file=stderr)
-    for siba,sibb in siblings:
-        if not siba in id_to_edges or not sibb in id_to_edges:
-            continue
-        x,y=id_to_edges[siba]
-        w,z =id_to_edges[sibb]
-        for u,v in [(x,y),(y,x)]:
-            sibmap[(u,v)]=(w,z)
-        for u,v in [(w,z),(z,w)]:
-            sibmap[(u,v)]=(x,y)
-    return sibmap
 
 
-
-
-
-def warm_start_decomposition(graphs,fam_bounds,families,siblings):
-    not_full_fams = du.getNotFullFamilies(fam_bounds,families)
-    create_max_weight(graphs,not_full_fams,siblings)
-    for tree_edge, ((child, parent), G) in enumerate(sorted(graphs.items())):
-        sibmap=create_sibmap(G,siblings[(child,parent)])
-        #cycle_greedy_partial_matching(G,sibmap)
-        heuristic_partial_matching(G,sibmap)
-        greedy_matching_completion(G,sibmap)
-
-
-def get_sup_families(G,not_full_fams,gnm):
-    nff_markers = dict()
-    #if the matching does not allow for a good node
-    for v,data in G.nodes(data=True):
-        f = du.nodeGetFamily(G,v)
-        if data['type']==du.VTYPE_CAP or du.get_genome(G,v)!=gnm:
-            continue
-        if data['id'][1][1]!=du.EXTR_TAIL:
-            continue
-        if f in not_full_fams[gnm]:
-            if f not in nff_markers:
-                nff_markers[f]=set()
-            nff_markers[f].add(v)
[…]

### 2025-01-31 example/run.sh
Add example, fix reading of local managers

@@ -1,0 +1,4 @@
+#!/usr/bin/env bash
+python3 ../scripts/spp_dcj.py -a 1 -m idmap.txt tree.txt adjacencies.txt > example.ilp
+python3 ../scripts/compute_initial.py -a 1 tree.txt adjacencies.txt idmap.txt example_init.sol
+gurobi_cl InputFile=example_init.sol example.ilp

### 2025-01-31 scripts/compute_initial.py

@@ -50,7 +50,7 @@ def main():
     du.fillFamilyBounds(families,fam_bounds)
     relationalDiagrams = du.constructRelationalDiagrams(speciesTree,
             adjacencies, telomeres, weights, genes, global_ext2id,fam_bounds=fam_bounds,
-            sep=args.separator)
+            sep=args.separator,loc_manager_tables=loc_tables)
     graphs = relationalDiagrams['graphs']
     siblings = relationalDiagrams['siblings']
     for G in graphs.values():

### 2025-01-31 scripts/compute_initial.py
Fix help messages of cl parsing

@@ -4,7 +4,7 @@ import data_utils as du
 def main():
     parser = ArgumentParser()
     parser.add_argument('tree', type=open,
-            help='phylogenetic tree as parent-child relation table')
+            help='phylogenetic tree as child->parent relation table (Important: children must be in first column.)')
     parser.add_argument('candidateAdjacencies', type=open,
             help='candidate adjacencies of the genomes in the phylogeny')
     parser.add_argument('idmap',help="Id map file (output of the main script)",type=open)

### 2025-01-31 scripts/spp_dcj.py

@@ -661,7 +661,7 @@ if __name__ == '__main__':
 
     parser = ArgumentParser(formatter_class=ADHF)
     parser.add_argument('tree', type=open,
-            help='phylogenetic tree as parent-child relation table')
+            help='phylogenetic tree as child->parent relation table (Important: children must be in first column.)')
     parser.add_argument('candidateAdjacencies', type=open,
             help='candidate adjacencies of the genomes in the phylogeny')
     parser.add_argument('-t', '--no_telomeres', action='store_true',

### 2025-01-31 example/run.sh
Fix sol parsing for adjacencies: Todo: unimog

@@ -1,4 +1,5 @@
 #!/usr/bin/env bash
 python3 ../scripts/spp_dcj.py -a 1 -m idmap.txt tree.txt adjacencies.txt > example.ilp
 python3 ../scripts/compute_initial.py -a 1 tree.txt adjacencies.txt idmap.txt example_init.sol
-gurobi_cl InputFile=example_init.sol example.ilp
+gurobi_cl InputFile=example_init.sol ResultFile=example.sol example.ilp 
+python3 ../scripts/sol2adjacencies.py example.sol idmap.txt 

### 2025-01-31 scripts/data_utils.py

@@ -421,7 +421,7 @@ def parse_edge_id(r):
 
 def parseSOL(data, idMap):
     """ SOL file parser """
-
+    raise NotImplementedError("Not implemented for this version of SPP-DCJ")
     obj_value = None
     adjacenciesList = dict()
     matchingList = set()
@@ -478,6 +478,41 @@ def parseSOL(data, idMap):
                 indelList[ext1[0]].append((ext1[1:], ext2[1:]))
     return adjacenciesList, indelList, weightsDict, sorted(matchingList), \
             obj_value, vars_
+
+def parseSOLAdj(data, idMap):
+    """ SOL file parser """
+    obj_value = None
+    adjacenciesList = dict()
+
+
+    vars_ = dict()
+
+    # objective value is stored in the following comment:
+    obj_txt = '# Objective value = '
+    for line in data:
+        if line.startswith(obj_txt):
+            obj_value = float(line[len(obj_txt):])
+            continue
+        #print(line,file=sys.stderr)
+        var_, val = line.split()
+        vars_[var_] = float(val)
+    #collect set adjacencies
+    for var_,val in vars_.items():
+        if abs(val - 1) > I_EPSILON:
+            continue
+        if var_.split(SEP)[0]=='a':
+            entries = var_.split(SEP)
+            #format aSEPedge
+            rest = SEP.join(entries[1::])
+            id1, id2, _ = parse_edge_id(rest)
+            ext1 = idMap[id1]
+            ext2 = idMap[id2]
+            if ext1[0] not in adjacenciesList:
+                adjacenciesList[ext1[0]] = list()
+            adj = (ext1[1], ext2[1])
+            adjacenciesList[ext1[0]].append(adj)
+    return adjacenciesList, obj_value
+
 
 #
 # CORE & CONVENIENCE FUNCTIONS

### 2025-01-31 scripts/sol2adjacencies.py

@@ -8,7 +8,7 @@ import re
 
 # import from own packages
 import data_utils as du
-
+from collections import defaultdict
 
 if __name__ == '__main__':
 
@@ -25,12 +25,12 @@ if __name__ == '__main__':
     #
 
     # id-to-extremity mapping
-    id2ext = dict()
-    for line in csv.reader(args.id_to_extremity_map, delimiter = '\t'):
-        if line:
-            id2ext[line[0]] = tuple(line[1:])
+    id2extr,_ = du.parse_id_file(args.id_to_extremity_map)#dict()
+    id2ext = dict([(str(v),k) for k,v in id2extr.items()])
+    #for line in csv.reader(args.id_to_extremity_map, delimiter = '\t'):
+    #    if line:
+    #        id2ext[line[0]] = tuple(line[1:])
 
-    adjacenciesList, _, weightsDict, _, _, _ = du.parseSOL(args.sol_file, id2ext)
-
+    adjacenciesList,_ = du.parseSOLAdj(args.sol_file, id2ext)
     # write adjacencies
-    du.writeAdjacencies(adjacenciesList, weightsDict, stdout)
+    du.writeAdjacencies(adjacenciesList,defaultdict(lambda: defaultdict(float)), stdout)

### 2025-01-31 scripts/sol2unimog.py

@@ -131,10 +131,10 @@ if __name__ == '__main__':
     #
 
     # id-to-extremity mapping
-    id2ext = dict()
-    for line in csv.reader(args.id_to_extremity_map, delimiter = '\t'):
-        if line:
-            id2ext[line[0]] = tuple(line[1:])
+    id2ext,_ = du.parse_id_file(args.id_to_extremity_map)
+    #for line in csv.reader(args.id_to_extremity_map, delimiter = '\t'):
+    #    if line:
+    #        id2ext[line[0]] = tuple(line[1:])
 
     adjacenciesList, indelList, _, matchingList, _, vars_ = \
             du.parseSOL(args.sol_file, id2ext)

### 2025-01-31 example/run.sh
Make example a bit nicer

@@ -2,4 +2,4 @@
 python3 ../scripts/spp_dcj.py -a 1 -m idmap.txt tree.txt adjacencies.txt > example.ilp
 python3 ../scripts/compute_initial.py -a 1 tree.txt adjacencies.txt idmap.txt example_init.sol
 gurobi_cl InputFile=example_init.sol ResultFile=example.sol example.ilp 
-python3 ../scripts/sol2adjacencies.py example.sol idmap.txt 
+python3 ../scripts/sol2adjacencies.py example.sol idmap.txt  > resolved_adjacencies.txt

### 2025-01-31 docs/manuscript/main.tex

@@
 \subsection{Preliminaries}
+%A \emph{(genomic) marker} $g := (g^\et, g^\eh)$ is a universally unique entity consisting of \emph{marker extremities} tail of $g$, denoted by $g^\et$, and head of $g$, denoted by $g^\eh$. 
+%A \emph{telomere} $t^\circ$ is a universally unique entity encompassing a single telomeric extremity denoted by ``$\circ$''. 
+%A genome is a triple $(M \cup \pseudocaps, A, \family)$, with $M$ being its marker set and $\pseudocaps$ its set of telomeres, $A \subset (M \cup \pseudocaps)^2$ the set of telomeres, and $\family: M \to \mathbb N$ a function indicating the family of each marker.
+%The set of adjacencies $A$ has the properties that \emph{(i)} for each extremity $\varepsilon \in \bigcup M \cup \pseudocaps$, there exists exactly one adjacency $a \in A$ that contains $\varepsilon$.
+%The adjacency set decomposes the set of markers and telomeres into linear or circular components, that we further call \emph{chromosomes}. 
+%Moreover, \emph{(ii)} each linear chromosome, formed by ordering markers and telomeres based on their adjacencies in $A$, starts and ends with a telomere. 
+%
+%A \emph{(genome) scaffold} is a quintuple $(M \cup \pseudocaps, A, \family, \multi, \w)$ with 
+%\begin{itemize}
+%    \item $M$, $\pseudocaps$, and $\family$ as defined in a genome, 
+%    \item the set of adjacencies $A$ satisfying that for each $g^\et \in \bigcup A$, there exists also extremity $g^\eh \in \bigcup A$ and vice versa, and each telomeric extremity is used only once, i.e., $\forall \{X, X'\} \subseteq A$, $X \cap X' \cap \pseudocaps = \emptyset$,
+%    \item copy number function $\multi(i) = [a, b]$ reporting for each family $i$ the permitted range $0 \leq a \leq b$ of its copy number, such that each family, defined as $F_i := \{g \in M \mid f(g) = i\}$, satisfies $|F_i| = b$, and
+%    \item adjacency weight function $\w: A \to \mathbb R$. % [-1, 1]$.
+%\end{itemize}
+%
+%Observe that a genome can give rise to a scaffold, with copy number $\multi$ and weight $\w$ functions arbitrarily defined, but the reverse does not hold true in general.
+%% not sure we need this:
+%% However, in the following, we regard genomes as scaffolds with tight copy numbers, that is, the copy number range of each family $i$ corresponds to a point that coincides with the number of its associated markers.
+%A genome $S' = (M', A', \family')$ is \emph{derived} from a scaffold $S = (M, A, \family, \multi, \w)$, or simply ``$S$-derived'', if \emph{(i)} $M' \subseteq M$, \emph{(ii)} $A' \subseteq A$, \emph{(iii)}
+%there exist no two markers $g, g' \in M'$ with $\family'(g) = \family'(g')$ and $\family(g) \neq \family(g')$, and \emph{(iv)} the original copy number constraints are satisfied, i.e., for each family $i$ and the set of markers associated with $i$ in scaffold $S$, i.e., $F_i = \{ g \in M' \mid \family(g) = i\}$, holds true that $|F_i| \in \multi(i)$.
+%We call a scaffold $S$ \emph{linearizable} if there exists an $S$-derived genome. 
+%In fact, many scaffolds are not linearizable. 
+%
+%A \emph{phylogeny} $\tree$ is a connected graph with nodes representing \emph{operational taxonomic units} (OTUs). 
+%Nodes with degree 1 are termed \emph{tips} of the phylogeny.
+%
+%General notation
+%\begin{itemize}
+%    \item 
+%    \item Weighted multigraph $G = (\vertices, \edges, \w)$ with edge weight function $\w: \edges \to \mathbb R$
+%\end{itemize}
+%
+%Paper-specific notation:
+%
+%\begin{itemize}
+%    \item Genomic marker 
+%    \item Furthermore, we use a function $\extf : \extremities \to \{\et, \eh, \circ\}$ to map extremities to their corresponding kind (tail, head or telomere). 
+%    \item 
+%We model family assignments of marker extremities as a function $\family: \extremities \to \mathbb N$ for which holds true that for any marker $g = \{g^\et, g^\eh\}$, $\family(g^\et) = \family(g^\eh)$.
+%Function 
+%$\minmulti_\X : \mathbb N \to \mathbb N$ reports the minimum multiplicity of a gene family in a given genome $\X$, while $\maxmulti_\X: \mathbb N \to \mathbb N$ reports its maximum multiplicity. 
+%    \item Multirelational diagram $\mrd$ and \emph{capping-free multi-relational diagram} (CFMRD) $\pcmrd$
+%\end{itemize}
+%
+%\begin{itemize}
+%    \item SPP algoirthms such as 
+%\end{itemize}
+%
+%\begin{problem}[Weighted CN-constrained degenerate DCJ indel distance]\label{prb:wdeg_dcj}
+%    Given a weighting scheme $\w : \extremities \times \extremities \to \mathbb R$, some $\alpha \in [0, 1]$, two linearizable degenerate genomes $\A, \B$ with copy number constraints \hl{XX}\todoD{add data structure} and family assignment $\family$, find $\A$-derived genome $\A'$, $\B$-derived genome $\B'$, and $\family$-derived $\{\A',\B'\}$-resolved family assignment $\family'$ that minimize the linear combination
+%    $$
+%    (1-\alpha) \cdot \sum_{X \in \A' \cup \B'} -\w(X) + \alpha \cdot \dist_\DCJid(A', B')\,.
+%    $$
+%\end{problem}
+%
+%\begin{problem}[SPP-DCJ]\label{prb:spp_dcj}
+%    Given a phylogeny $\tree$ and a set of linearizable degenerate genomes $\A_1, \ldots, \A_k$ corresponding to the node set $V(\tree) = \{\A_1, \ldots, \A_k\}$, find genomes $\A'_1 \subseteq \A_1, \ldots, \A'_k \allowbreak \subseteq \A_k$ that minimize the sum of weighted degenerate DCJ indel distances along the edges of $\tree$.
+%\end{problem}
+
+For the purposes of this work, we use the longstanding abstraction to describe genomes as sequences of oriented markers.
 A \emph{(genomic) marker} $g := (g^\et, g^\eh)$ is a universally unique entity consisting of \emph{marker extremities} tail of $g$, denoted by $g^\et$, and head of $g$, denoted by $g^\eh$.
+
+The structure of a genome can be described via its adjacencies.
+An adjacency $\{m^x,n^y\}$ describes that markers $m,n$ are on the same chromosome oriented, such that extremities $m^x,n^y$ are directly neighboring.
+For ease of notation we also write $m^xn^y$ for an adjacency.
+Note that it may also be read in the other direction, i.e.
+$n^ym^x$ is equivalent to $m^xn^y$.
+
+For ease of notation, we aim for each extremity to be part of an adjacency. In order to accomplish this, we use additional extremities modeling the ends of linear chromosomes, called \emph{telomeres}.
 A \emph{telomere} $t^\circ$ is a universally unique entity encompassing a single telomeric extremity denoted by ``$\circ$''.
-A genome is a triple $(M \cup \pseudocaps, A, \family)$, with $M$ being its marker set and $\pseudocaps$ its set of telomeres, $A \subset (M \cup \pseudocaps)^2$ the set of telomeres, and $\family: M \to \mathbb N$ a function indicating the family of each marker.
-The set of adjacencies $A$ has the properties that \emph{(i)} for each extremity $\varepsilon \in \bigcup M \cup \pseudocaps$, there exists exactly one adjacency $a \in A$ that contains $\varepsilon$.
-The adjacency set decomposes the set of markers and telomeres into linear or circular components, that we further call \emph{chromosomes}. 
-Moreover, \emph{(ii)} each linear chromosome, formed by ordering markers and telomeres based on their adjacencies in $A$, starts and ends with a telomere.
+A genome can then be described as a graph as follows.
 
-A \emph{(genome) scaffold} is a quintuple $(M \cup \pseudocaps, A, \family, \multi, \w)$ with 
-\begin{itemize}
-    \item $M$, $\pseudocaps$, and $\family$ as defined in a genome, 
-    \item the set of adjacencies $A$ satisfying that for each $g^\et \in \bigcup A$, there exists also extremity $g^\eh \in \bigcup A$ and vice versa, and each telomeric extremity is used only once, i.e., $\forall \{X, X'\} \subseteq A$, $X \cap X' \cap \pseudocaps = \emptyset$,
-    \item copy number function $\multi(i) = [a, b]$ reporting for each family $i$ the permitted range $0 \leq a \leq b$ of its copy number, such that each family, defined as $F_i := \{g \in M \mid f(g) = i\}$, satisfies $|F_i| = b$, and
-    \item adjacency weight function $\w: A \to \mathbb R$. % [-1, 1]$.
-\end{itemize}
+\begin{definition}\label{def:genome}
+A genome $\A$ is a graph with vertices $\extremities(\A)\cup\telomeres (\A)$, namely its extremities $\extremities(\A)$ and  telomeres $\telomeres (\A)$.
+The set of edges is $\markers(\A)\cup\adjacencies(\A)$, namely its markers $\markers(\A)$ and adjacencies $\adjacencies(\A)$.
+This graph fulfills the following properties:
+\begin{enumerate}
+	\item\label{itm:markers} $\markers (\A)$ is a perfect matching on $\extremities(\A)$ and
+	\item $\adjacencies(\A)$ is a perfect matching on $\extremities(\A)\cup\telomeres (\A)$.
+\end{enumerate}
+\end{definition}
 
-Observe that a genome can give rise to a scaffold, with copy number $\multi$ and weight $\w$ functions arbitrarily defined, but the reverse does not hold true in general.
-% not sure we need this:
-% However, in the following, we regard genomes as scaffolds with tight copy numbers, that is, the copy number range of each family $i$ corresponds to a point that coincides with the number of its associated markers.
-A genome $S' = (M', A', \family')$ is \emph{derived} from a scaffold $S = (M, A, \family, \multi, \w)$, or simply ``$S$-derived'', if \emph{(i)} $M' \subseteq M$, \emph{(ii)} $A' \subseteq A$, \emph{(iii)}
-there exist no two markers $g, g' \in M'$ with $\family'(g) = \family'(g')$ and $\family(g) \neq \family(g')$, and \emph{(iv)} the original copy number constraints are satisfied, i.e., for each family $i$ and the set of markers associated with $i$ in scaffold $S$, i.e., $F_i = \{ g \in M' \mid \family(g) = i\}$, holds true that $|F_i| \in \multi(i)$.
-We call a scaffold $S$ \emph{linearizable} if there exists an $S$-derived genome. 
-In fact, many scaffolds are not linearizable.
+A degenerate genome $\D$ is a graph $(\extremities(\D)\cup\telomeres (\D),\markers(\D)\cup\adjacencies(\D))$ that fulfills Requirement~\ref{itm:markers} of Definition~\ref{def:genome}.
+A \emph{\gls*{linr}} of a degenerate genome $\D$ is a genome $\A$, such that $\extremities(\A)=\extremities(\D)$, $\telomeres (\A) \subset \telomeres (\D)$, $\markers(\A)=\markers(\D)$ and $\adjacencies(\A)\subset\adjacencies(\D)$.
+If such a linearization exists, we call $\D$ \emph{linearizable}.
 
-A \emph{phylogeny} $\tree$ is a connected graph with nodes representing \emph{operational taxonomic units} (OTUs). 
-Nodes with degree 1 are termed \emph{tips} of the phylogeny.
+\begin{definition}
+	DCJ-indel distance
+\end{definition}
 
-General notation
-\begin{itemize}
-    \item 
-    \item Weighted multigraph $G = (\vertices, \edges, \w)$ with edge weight function $\w: \edges \to \mathbb R$
-\end{itemize}
-
-Paper-specific notation:
-
-\begin{itemize}
-    \item Genomic marker 
-    \item Furthermore, we use a function $\extf : \extremities \to \{\et, \eh, \circ\}$ to map extremities to their corresponding kind (tail, head or telomere). 
-    \item 
-We model family assignments of marker extremities as a function $\family: \extremities \to \mathbb N$ for which holds true that for any marker $g = \{g^\et, g^\eh\}$, $\family(g^\et) = \family(g^\eh)$.
-Function 
-$\minmulti_\X : \mathbb N \to \mathbb N$ reports the minimum multiplicity of a gene family in a given genome $\X$, while $\maxmulti_\X: \mathbb N \to \mathbb N$ reports its maximum multiplicity. 
-    \item Multirelational diagram $\mrd$ and \emph{capping-free multi-relational diagram} (CFMRD) $\pcmrd$
-\end{itemize}
-
-\begin{itemize}
-    \item SPP algoirthms such as 
-\end{itemize}
-
-\begin{problem}[Weighted CN-constrained degenerate DCJ indel distance]\label{prb:wdeg_dcj}
-    Given a weighting scheme $\w : \extremities \times \extremities \to \mathbb R$, some $\alpha \in [0, 1]$, two linearizable degenerate genomes $\A, \B$ with copy number constraints \hl{XX}\todoD{add data structure} and family assignment $\family$, find $\A$-derived genome $\A'$, $\B$-derived genome $\B'$, and $\family$-derived $\{\A',\B'\}$-resolved family assignment $\family'$ that minimize the linear combination
-    $$
-    (1-\alpha) \cdot \sum_{X \in \A' \cup \B'} -\w(X) + \alpha \cdot \dist_\DCJid(A', B')\,.
-    $$
-\end{problem}
-
-\begin{problem}[SPP-DCJ]\label{prb:spp_dcj}
-    Given a phylogeny $\tree$ and a set of linearizable degenerate genomes $\A_1, \ldots, \A_k$ corresponding to the node set $V(\tree) = \{\A_1, \ldots, \A_k\}$, find genomes $\A'_1 \subseteq \A_1, \ldots, \A'_k \allowbreak \subseteq \A_k$ that minimize the sum of weighted degenerate DCJ indel distances along the edges of $\tree$.
-\end{problem}
-
+\begin{definition}[Weighted Small Parsimony Linearization Problem]
+	Given a phylogeny $\Gamma$ and a weighting function for adjacencies $w$ find a linearization $\Lr_i$ for each (degenerate) genome $\D_i$ in $\Gamma$, such that
+	\begin{equation}
+		\alpha \sum_{(\D_i,\D_k)\in \Gamma} \ddcjid (\Lr_i,\Lr_k) + (\alpha -1) \sum_{\D_i\in \Gamma}\sum_{ab\in\adjacencies(\Lr_i)}w(ab).
+	\end{equation}
+	is minimized.
+\end{definition}
 \section{A new Method}
